---
title: "RPC-PI Experiment - Win Rate Analysis"
author: ""
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    theme: sandstone
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)

options(knitr.kable.NA = "-")

required_packages <- c(
  "tidyverse",
  "readr",
  "ggplot2",
  "scales",
  "stringr",
  "forcats",
  "fs",
  "here",
  "knitr",
  "jsonlite"
)

missing_packages <- required_packages[!required_packages %in% rownames(installed.packages())]
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

library(tidyverse)
library(readr)
library(ggplot2)
library(scales)
library(stringr)
library(forcats)
library(fs)
library(here)
library(knitr)
library(jsonlite)
```

# Objective

This R Markdown focuses on **judged preference / Win Rate**.

It is different from accuracy and ASV:

- accuracy asks whether the model predicted the correct task label;
- ASV asks whether the attack reached the adversarial target;
- Win Rate asks whether a judge preferred the defended model output over the C0 reference output.

The key metric is:

\[
AdjustedWinRate = \frac{wins + 0.5 \cdot ties}{valid\_judgments}
\]

Interpretation:

- `adjusted_win_rate > 0.50`: the defended scenario is preferred over C0 on average;
- `adjusted_win_rate = 0.50`: parity with C0;
- `adjusted_win_rate < 0.50`: C0 is preferred on average.

In this experiment, the win-rate artifacts compare each defended scenario against `C0 - Base model` on `test_clean`.

# Paths and configuration

```{r paths}
DATA_DIR <- here::here("data", "results")
PAIRWISE_DIR <- file.path(DATA_DIR, "pairwise_metrics", "full")

OUT_TABLE_DIR <- here::here("tables", "win_rate_analysis")
OUT_FIG_DIR <- here::here("figures", "win_rate_analysis")

fs::dir_create(OUT_TABLE_DIR)
fs::dir_create(OUT_FIG_DIR)

if (!dir.exists(DATA_DIR)) {
  stop(
    "Could not find data/results/. Extract the ZIP first with: ",
    "unzip('results.zip', exdir = 'data')"
  )
}

if (!dir.exists(PAIRWISE_DIR)) {
  stop("Could not find pairwise metrics directory: ", PAIRWISE_DIR)
}
```

# Helper functions

```{r helpers}
read_first_existing <- function(paths, required = TRUE) {
  existing <- paths[file.exists(paths)]

  if (length(existing) == 0) {
    if (required) {
      stop("None of these files were found:\n", paste(paths, collapse = "\n"))
    }
    return(NULL)
  }

  readr::read_csv(existing[[1]], show_col_types = FALSE)
}

read_jsonl_file <- function(path, bad_lines_filename = "00_bad_jsonl_lines.csv") {
  if (!file.exists(path)) {
    return(NULL)
  }

  # Robust JSONL reader:
  # jsonlite::stream_in() stops at the first malformed/truncated line.
  # Here we parse line by line, keep valid records, and save malformed lines
  # for inspection instead of stopping the whole report.
  lines <- readLines(path, warn = FALSE, encoding = "UTF-8")

  if (length(lines) == 0) {
    return(tibble())
  }

  parsed <- vector("list", length(lines))
  bad_lines <- list()

  for (i in seq_along(lines)) {
    line <- lines[[i]]

    if (is.na(line) || trimws(line) == "") {
      next
    }

    obj <- tryCatch(
      jsonlite::fromJSON(line, flatten = TRUE),
      error = function(e) {
        bad_lines[[length(bad_lines) + 1]] <<- tibble(
          line_number = i,
          error = conditionMessage(e),
          line_prefix = substr(line, 1, 250)
        )
        NULL
      }
    )

    if (!is.null(obj)) {
      parsed[[i]] <- tibble::as_tibble(obj)
    }
  }

  parsed <- parsed[!vapply(parsed, is.null, logical(1))]

  bad_lines_df <- dplyr::bind_rows(bad_lines)

  if (nrow(bad_lines_df) > 0) {
    readr::write_csv(
      bad_lines_df,
      file.path(OUT_TABLE_DIR, bad_lines_filename)
    )

    message(
      "Skipped ",
      nrow(bad_lines_df),
      " malformed JSONL line(s). See ",
      file.path(OUT_TABLE_DIR, bad_lines_filename)
    )
  }

  if (length(parsed) == 0) {
    return(tibble())
  }

  dplyr::bind_rows(parsed)
}

scenario_order <- c(
  "c0_base",
  "c1_struq_format_only",
  "c2_struq_sft",
  "c3_secalign_dpo",
  "c4_ih_sft"
)

scenario_labels <- tibble(
  scenario_id = scenario_order,
  scenario_short = c("C0", "C1", "C2", "C3", "C4"),
  scenario_label = c(
    "C0 - Base model",
    "C1 - StruQ format-only",
    "C2 - StruQ-like SFT",
    "C3 - SecAlign-like DPO",
    "C4 - Instruction-Hierarchy-like SFT"
  )
)

scenario_palette <- c(
  C0 = "#7F7F7F",
  C1 = "#D55E00",
  C2 = "#0072B2",
  C3 = "#CC79A7",
  C4 = "#009E73"
)

sort_scenarios <- function(df) {
  df |>
    mutate(
      scenario_id = factor(as.character(scenario_id), levels = scenario_order),
      scenario_short = factor(as.character(scenario_short), levels = c("C0", "C1", "C2", "C3", "C4"))
    ) |>
    arrange(scenario_id)
}

add_scenario_info <- function(df, id_col = "scenario_id", label_col = NULL) {
  out <- df |>
    mutate(scenario_id = as.character(.data[[id_col]]))

  if (!is.null(label_col) && label_col %in% names(out)) {
    out <- out |>
      mutate(scenario_label = as.character(.data[[label_col]]))
  }

  out <- out |>
    left_join(scenario_labels, by = "scenario_id", suffix = c("", "_lookup"))

  if (!("scenario_label" %in% names(out))) {
    out$scenario_label <- out$scenario_id
  }

  if ("scenario_label_lookup" %in% names(out)) {
    out <- out |>
      mutate(
        scenario_label = coalesce(
          as.character(scenario_label),
          as.character(scenario_label_lookup)
        )
      )
  }

  if (!("scenario_short" %in% names(out))) {
    out$scenario_short <- NA_character_
  }

  if ("scenario_short_lookup" %in% names(out)) {
    out <- out |>
      mutate(
        scenario_short = coalesce(
          as.character(scenario_short),
          as.character(scenario_short_lookup)
        )
      )
  }

  out |>
    mutate(
      scenario_short = ifelse(
        is.na(scenario_short),
        recode(
          as.character(scenario_id),
          c0_base = "C0",
          c1_struq_format_only = "C1",
          c2_struq_sft = "C2",
          c3_secalign_dpo = "C3",
          c4_ih_sft = "C4",
          .default = as.character(scenario_id)
        ),
        scenario_short
      )
    ) |>
    select(-any_of(c("scenario_label_lookup", "scenario_short_lookup"))) |>
    sort_scenarios()
}

to_logical_safe <- function(x) {
  if (is.logical(x)) return(x)

  x_chr <- tolower(as.character(x))

  case_when(
    x_chr %in% c("true", "t", "1", "yes") ~ TRUE,
    x_chr %in% c("false", "f", "0", "no") ~ FALSE,
    TRUE ~ NA
  )
}

safe_sd <- function(x) {
  if (sum(!is.na(x)) <= 1) return(NA_real_)
  sd(x, na.rm = TRUE)
}

safe_sem <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) return(NA_real_)
  sd(x, na.rm = TRUE) / sqrt(n)
}

safe_tcrit <- function(n, confidence = 0.95) {
  if (is.na(n) || n <= 1) return(NA_real_)
  alpha <- 1 - confidence
  qt(1 - alpha / 2, df = n - 1)
}

wilson_ci <- function(k, n, conf.level = 0.95) {
  if (is.na(k) || is.na(n) || n <= 0) {
    return(tibble(estimate = NA_real_, ci_low = NA_real_, ci_high = NA_real_))
  }

  z <- qnorm(1 - (1 - conf.level) / 2)
  p <- k / n
  denom <- 1 + z^2 / n
  center <- (p + z^2 / (2 * n)) / denom
  half_width <- z * sqrt((p * (1 - p) / n) + (z^2 / (4 * n^2))) / denom

  tibble(
    estimate = p,
    ci_low = max(0, center - half_width),
    ci_high = min(1, center + half_width)
  )
}

safe_binom_p <- function(x, n, p = 0.5, alternative = "greater") {
  if (is.na(x) || is.na(n) || n <= 0 || x < 0 || x > n) {
    return(NA_real_)
  }

  tryCatch(
    binom.test(as.integer(x), as.integer(n), p = p, alternative = alternative)$p.value,
    error = function(e) NA_real_
  )
}

fmt_pct <- function(x, accuracy = 0.1) {
  scales::percent(x, accuracy = accuracy)
}

base_theme <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      legend.position = "bottom",
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(color = "grey30"),
      panel.grid.minor = element_blank(),
      axis.title = element_text(face = "bold")
    )
}

write_win_csv <- function(df, filename) {
  readr::write_csv(df, file.path(OUT_TABLE_DIR, filename))
  invisible(df)
}

save_win_figure <- function(plot, filename, width = 10, height = 6) {
  ggsave(
    file.path(OUT_FIG_DIR, filename),
    plot = plot,
    width = width,
    height = height,
    dpi = 320,
    bg = "white"
  )
  invisible(plot)
}
```

# Load win-rate metrics

```{r load-win-metrics}
win_metrics_path <- file.path(PAIRWISE_DIR, "win_rate_metrics.csv")

win_metrics_raw <- read_first_existing(c(win_metrics_path))

win_metrics <- win_metrics_raw |>
  rename(
    scenario_id = defended_scenario_id,
    seed = defended_seed
  ) |>
  add_scenario_info(id_col = "scenario_id", label_col = "defended_scenario_label") |>
  mutate(
    seed = as.integer(seed),
    eval_split = as.character(eval_split),
    wins = as.integer(wins),
    losses = as.integer(losses),
    ties = as.integer(ties),
    valid_judgments = as.integer(valid_judgments),
    n_judgments = as.integer(n_judgments),
    win_rate_recomputed = if_else(valid_judgments > 0, wins / valid_judgments, NA_real_),
    loss_rate_recomputed = if_else(valid_judgments > 0, losses / valid_judgments, NA_real_),
    tie_rate_recomputed = if_else(valid_judgments > 0, ties / valid_judgments, NA_real_),
    adjusted_win_rate_recomputed = if_else(
      valid_judgments > 0,
      (wins + 0.5 * ties) / valid_judgments,
      NA_real_
    ),
    non_tie_judgments = wins + losses,
    non_tie_win_rate = if_else(non_tie_judgments > 0, wins / non_tie_judgments, NA_real_),
    net_win_rate = win_rate - loss_rate,
    adjusted_win_rate_delta_from_parity = adjusted_win_rate - 0.5,
    preference_direction = case_when(
      adjusted_win_rate > 0.5 ~ "Defended preferred",
      adjusted_win_rate < 0.5 ~ "C0 preferred",
      TRUE ~ "Parity"
    )
  ) |>
  sort_scenarios()

write_win_csv(win_metrics, "01_win_rate_metrics_enriched.csv")

win_metrics |>
  select(
    scenario_short,
    seed,
    eval_split,
    n_judgments,
    valid_judgments,
    wins,
    losses,
    ties,
    win_rate,
    loss_rate,
    tie_rate,
    adjusted_win_rate,
    non_tie_win_rate,
    net_win_rate,
    invalid_judgment_rate
  ) |>
  mutate(
    across(
      c(
        win_rate,
        loss_rate,
        tie_rate,
        adjusted_win_rate,
        non_tie_win_rate,
        net_win_rate,
        invalid_judgment_rate
      ),
      ~ fmt_pct(.x)
    )
  ) |>
  knitr::kable(
    caption = "Table 1 - Win-rate metrics by scenario and seed."
  )
```

# Validate recomputation

```{r validate-recomputed-metrics}
win_metric_validation <- win_metrics |>
  transmute(
    scenario_short,
    seed,
    eval_split,
    win_rate_diff = win_rate_recomputed - win_rate,
    loss_rate_diff = loss_rate_recomputed - loss_rate,
    tie_rate_diff = tie_rate_recomputed - tie_rate,
    adjusted_win_rate_diff = adjusted_win_rate_recomputed - adjusted_win_rate
  )

write_win_csv(win_metric_validation, "02_win_rate_recomputation_validation.csv")

win_metric_validation |>
  summarise(
    max_abs_win_rate_diff = max(abs(win_rate_diff), na.rm = TRUE),
    max_abs_loss_rate_diff = max(abs(loss_rate_diff), na.rm = TRUE),
    max_abs_tie_rate_diff = max(abs(tie_rate_diff), na.rm = TRUE),
    max_abs_adjusted_win_rate_diff = max(abs(adjusted_win_rate_diff), na.rm = TRUE)
  ) |>
  knitr::kable(
    caption = "Table 2 - Maximum absolute recomputation differences."
  )
```

# Scenario-level summary across seeds

```{r scenario-summary}
scenario_win_summary <- win_metrics |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    eval_split
  ) |>
  summarise(
    n_seeds = n_distinct(seed),
    seeds = paste(sort(unique(seed)), collapse = ", "),
    total_judgments = sum(n_judgments, na.rm = TRUE),
    total_valid_judgments = sum(valid_judgments, na.rm = TRUE),
    total_wins = sum(wins, na.rm = TRUE),
    total_losses = sum(losses, na.rm = TRUE),
    total_ties = sum(ties, na.rm = TRUE),

    mean_win_rate = mean(win_rate, na.rm = TRUE),
    mean_loss_rate = mean(loss_rate, na.rm = TRUE),
    mean_tie_rate = mean(tie_rate, na.rm = TRUE),
    mean_adjusted_win_rate = mean(adjusted_win_rate, na.rm = TRUE),
    sd_adjusted_win_rate = safe_sd(adjusted_win_rate),
    sem_adjusted_win_rate = safe_sem(adjusted_win_rate),
    tcrit_95 = safe_tcrit(n_distinct(seed)),
    ci95_low_adjusted_win_rate = mean_adjusted_win_rate - tcrit_95 * sem_adjusted_win_rate,
    ci95_high_adjusted_win_rate = mean_adjusted_win_rate + tcrit_95 * sem_adjusted_win_rate,
    mean_non_tie_win_rate = mean(non_tie_win_rate, na.rm = TRUE),
    mean_net_win_rate = mean(net_win_rate, na.rm = TRUE),
    mean_invalid_judgment_rate = mean(invalid_judgment_rate, na.rm = TRUE),

    weighted_win_rate = if_else(total_valid_judgments > 0, total_wins / total_valid_judgments, NA_real_),
    weighted_loss_rate = if_else(total_valid_judgments > 0, total_losses / total_valid_judgments, NA_real_),
    weighted_tie_rate = if_else(total_valid_judgments > 0, total_ties / total_valid_judgments, NA_real_),
    weighted_adjusted_win_rate = if_else(
      total_valid_judgments > 0,
      (total_wins + 0.5 * total_ties) / total_valid_judgments,
      NA_real_
    ),
    weighted_non_tie_win_rate = if_else(
      total_wins + total_losses > 0,
      total_wins / (total_wins + total_losses),
      NA_real_
    ),
    weighted_net_win_rate = weighted_win_rate - weighted_loss_rate,
    .groups = "drop"
  ) |>
  mutate(
    adjusted_win_rate_delta_from_parity = weighted_adjusted_win_rate - 0.5,
    preference_direction = case_when(
      weighted_adjusted_win_rate > 0.5 ~ "Defended preferred",
      weighted_adjusted_win_rate < 0.5 ~ "C0 preferred",
      TRUE ~ "Parity"
    )
  ) |>
  sort_scenarios()

write_win_csv(scenario_win_summary, "03_scenario_win_rate_summary.csv")

scenario_win_summary |>
  select(
    scenario_short,
    eval_split,
    n_seeds,
    seeds,
    total_valid_judgments,
    total_wins,
    total_losses,
    total_ties,
    weighted_win_rate,
    weighted_loss_rate,
    weighted_tie_rate,
    weighted_adjusted_win_rate,
    weighted_non_tie_win_rate,
    weighted_net_win_rate,
    mean_invalid_judgment_rate,
    preference_direction
  ) |>
  mutate(
    across(
      c(
        weighted_win_rate,
        weighted_loss_rate,
        weighted_tie_rate,
        weighted_adjusted_win_rate,
        weighted_non_tie_win_rate,
        weighted_net_win_rate,
        mean_invalid_judgment_rate
      ),
      ~ fmt_pct(.x)
    )
  ) |>
  knitr::kable(
    caption = "Table 3 - Scenario-level win-rate summary."
  )
```

# Judge-level rows

```{r load-judgments}
judgments_path <- file.path(PAIRWISE_DIR, "win_rate_judgments.jsonl")
judgments <- read_jsonl_file(judgments_path, bad_lines_filename = "00_bad_win_rate_judgment_lines.csv")

if (is.null(judgments)) {
  cat("win_rate_judgments.jsonl not found. Row-level sections will be skipped.")
} else {
  judgments <- judgments |>
    rename(
      scenario_id = defended_scenario_id,
      seed = defended_seed
    ) |>
    add_scenario_info(id_col = "scenario_id", label_col = "defended_scenario_label") |>
    mutate(
      seed = as.integer(seed),
      eval_split = as.character(eval_split),
      task_name = as.character(task_name),
      attack_type = ifelse(is.na(attack_type) | attack_type == "", "clean", as.character(attack_type)),
      judge_decision = as.character(judge_decision),
      judge_valid_decision = to_logical_safe(judge_valid_decision),
      reference_wins = to_logical_safe(reference_wins),
      defended_wins = to_logical_safe(defended_wins),
      tie = to_logical_safe(tie),
      valid_preference_score = case_when(
        judge_valid_decision & defended_wins ~ 1,
        judge_valid_decision & tie ~ 0.5,
        judge_valid_decision & reference_wins ~ 0,
        TRUE ~ NA_real_
      ),
      defended_win_int = as.integer(judge_valid_decision & defended_wins),
      reference_win_int = as.integer(judge_valid_decision & reference_wins),
      tie_int = as.integer(judge_valid_decision & tie),
      valid_judgment_int = as.integer(judge_valid_decision)
    ) |>
    sort_scenarios()

  write_win_csv(
    judgments |>
      select(
        eval_split,
        example_id,
        task_name,
        attack_type,
        reference_scenario_id,
        reference_seed,
        scenario_id,
        scenario_short,
        seed,
        expected_answer,
        attack_target,
        judge_model_id,
        judge_decision,
        judge_valid_decision,
        reference_wins,
        defended_wins,
        tie,
        valid_preference_score
      ),
    "04_win_rate_judgments_compact.csv"
  )

  judgments |>
    count(scenario_short, seed, eval_split, judge_decision, name = "n") |>
    knitr::kable(
      caption = "Table 4 - Judge decision distribution by scenario and seed."
    )
}
```

# Recompute metrics from judge-level rows

```{r recompute-from-judgments}
if (exists("judgments") && !is.null(judgments)) {
  recomputed_from_judgments <- judgments |>
    group_by(
      scenario_id,
      scenario_short,
      scenario_label,
      seed,
      eval_split
    ) |>
    summarise(
      n_judgments = n(),
      valid_judgments = sum(valid_judgment_int, na.rm = TRUE),
      wins = sum(defended_win_int, na.rm = TRUE),
      losses = sum(reference_win_int, na.rm = TRUE),
      ties = sum(tie_int, na.rm = TRUE),
      valid_judgment_rate = valid_judgments / n_judgments,
      win_rate = wins / valid_judgments,
      loss_rate = losses / valid_judgments,
      tie_rate = ties / valid_judgments,
      adjusted_win_rate = (wins + 0.5 * ties) / valid_judgments,
      invalid_judgment_rate = 1 - valid_judgment_rate,
      non_tie_judgments = wins + losses,
      non_tie_win_rate = if_else(non_tie_judgments > 0, wins / non_tie_judgments, NA_real_),
      net_win_rate = win_rate - loss_rate,
      .groups = "drop"
    ) |>
    sort_scenarios()

  write_win_csv(recomputed_from_judgments, "05_recomputed_win_rate_from_judgments.csv")

  comparison_with_metrics <- recomputed_from_judgments |>
    select(
      scenario_id,
      seed,
      eval_split,
      adjusted_win_rate_recomputed_rows = adjusted_win_rate,
      win_rate_recomputed_rows = win_rate,
      loss_rate_recomputed_rows = loss_rate,
      tie_rate_recomputed_rows = tie_rate
    ) |>
    left_join(
      win_metrics |>
        select(
          scenario_id,
          seed,
          eval_split,
          adjusted_win_rate,
          win_rate,
          loss_rate,
          tie_rate
        ),
      by = c("scenario_id", "seed", "eval_split")
    ) |>
    mutate(
      adjusted_diff = adjusted_win_rate_recomputed_rows - adjusted_win_rate,
      win_rate_diff = win_rate_recomputed_rows - win_rate,
      loss_rate_diff = loss_rate_recomputed_rows - loss_rate,
      tie_rate_diff = tie_rate_recomputed_rows - tie_rate
    )

  write_win_csv(comparison_with_metrics, "06_judgment_recomputation_vs_metrics.csv")

  comparison_with_metrics |>
    summarise(
      max_abs_adjusted_diff = max(abs(adjusted_diff), na.rm = TRUE),
      max_abs_win_rate_diff = max(abs(win_rate_diff), na.rm = TRUE),
      max_abs_loss_rate_diff = max(abs(loss_rate_diff), na.rm = TRUE),
      max_abs_tie_rate_diff = max(abs(tie_rate_diff), na.rm = TRUE)
    ) |>
    knitr::kable(
      caption = "Table 5 - Validation of judgment-level recomputation."
    )
}
```

# Wilson intervals and sign test for non-tie preferences

The adjusted win rate includes ties as half a win. For direct preference among non-tie decisions, we also analyze:

\[
NonTieWinRate = \frac{wins}{wins + losses}
\]

This asks: when the judge did not tie, how often did the defended scenario beat C0?

```{r non-tie-wilson-sign-test}
non_tie_inference <- win_metrics |>
  mutate(
    non_tie_judgments = wins + losses
  ) |>
  rowwise() |>
  mutate(
    wilson = list(wilson_ci(wins, non_tie_judgments, conf.level = 0.95))
  ) |>
  unnest(wilson) |>
  ungroup() |>
  rename(
    non_tie_win_rate_wilson = estimate,
    non_tie_win_rate_ci_low = ci_low,
    non_tie_win_rate_ci_high = ci_high
  ) |>
  mutate(
    sign_test_p_value_greater_than_parity = purrr::map2_dbl(
      wins,
      non_tie_judgments,
      ~ safe_binom_p(.x, .y, p = 0.5, alternative = "greater")
    ),
    sign_test_p_value_less_than_parity = purrr::map2_dbl(
      wins,
      non_tie_judgments,
      ~ safe_binom_p(.x, .y, p = 0.5, alternative = "less")
    ),
    significant_defended_preferred_05 = !is.na(sign_test_p_value_greater_than_parity) &
      sign_test_p_value_greater_than_parity < 0.05,
    significant_c0_preferred_05 = !is.na(sign_test_p_value_less_than_parity) &
      sign_test_p_value_less_than_parity < 0.05
  ) |>
  sort_scenarios()

write_win_csv(non_tie_inference, "07_non_tie_preference_wilson_and_sign_test.csv")

non_tie_inference |>
  select(
    scenario_short,
    seed,
    eval_split,
    wins,
    losses,
    ties,
    non_tie_judgments,
    non_tie_win_rate,
    non_tie_win_rate_ci_low,
    non_tie_win_rate_ci_high,
    sign_test_p_value_greater_than_parity,
    sign_test_p_value_less_than_parity
  ) |>
  mutate(
    across(
      c(non_tie_win_rate, non_tie_win_rate_ci_low, non_tie_win_rate_ci_high),
      ~ fmt_pct(.x)
    ),
    sign_test_p_value_greater_than_parity = scales::pvalue(sign_test_p_value_greater_than_parity, accuracy = 0.001),
    sign_test_p_value_less_than_parity = scales::pvalue(sign_test_p_value_less_than_parity, accuracy = 0.001)
  ) |>
  knitr::kable(
    caption = "Table 6 - Non-tie preference Wilson intervals and sign tests."
  )
```

# Task-level win-rate analysis

```{r task-level-analysis}
if (exists("judgments") && !is.null(judgments)) {
  win_rate_by_task <- judgments |>
    group_by(
      scenario_id,
      scenario_short,
      scenario_label,
      seed,
      eval_split,
      task_name
    ) |>
    summarise(
      n_judgments = n(),
      valid_judgments = sum(valid_judgment_int, na.rm = TRUE),
      wins = sum(defended_win_int, na.rm = TRUE),
      losses = sum(reference_win_int, na.rm = TRUE),
      ties = sum(tie_int, na.rm = TRUE),
      win_rate = wins / valid_judgments,
      loss_rate = losses / valid_judgments,
      tie_rate = ties / valid_judgments,
      adjusted_win_rate = (wins + 0.5 * ties) / valid_judgments,
      non_tie_judgments = wins + losses,
      non_tie_win_rate = if_else(non_tie_judgments > 0, wins / non_tie_judgments, NA_real_),
      net_win_rate = win_rate - loss_rate,
      invalid_judgment_rate = 1 - valid_judgments / n_judgments,
      .groups = "drop"
    ) |>
    sort_scenarios()

  task_summary <- win_rate_by_task |>
    group_by(
      scenario_id,
      scenario_short,
      scenario_label,
      eval_split,
      task_name
    ) |>
    summarise(
      n_seeds = n_distinct(seed),
      total_valid_judgments = sum(valid_judgments, na.rm = TRUE),
      total_wins = sum(wins, na.rm = TRUE),
      total_losses = sum(losses, na.rm = TRUE),
      total_ties = sum(ties, na.rm = TRUE),
      weighted_win_rate = total_wins / total_valid_judgments,
      weighted_loss_rate = total_losses / total_valid_judgments,
      weighted_tie_rate = total_ties / total_valid_judgments,
      weighted_adjusted_win_rate = (total_wins + 0.5 * total_ties) / total_valid_judgments,
      weighted_net_win_rate = weighted_win_rate - weighted_loss_rate,
      mean_adjusted_win_rate = mean(adjusted_win_rate, na.rm = TRUE),
      sd_adjusted_win_rate = safe_sd(adjusted_win_rate),
      .groups = "drop"
    ) |>
    sort_scenarios()

  write_win_csv(win_rate_by_task, "08_win_rate_by_task_seed.csv")
  write_win_csv(task_summary, "09_win_rate_by_task_summary.csv")

  task_summary |>
    select(
      scenario_short,
      task_name,
      total_valid_judgments,
      total_wins,
      total_losses,
      total_ties,
      weighted_adjusted_win_rate,
      weighted_net_win_rate
    ) |>
    mutate(
      weighted_adjusted_win_rate = fmt_pct(weighted_adjusted_win_rate),
      weighted_net_win_rate = fmt_pct(weighted_net_win_rate)
    ) |>
    arrange(scenario_short, task_name) |>
    knitr::kable(
      caption = "Table 7 - Task-level adjusted win rate summary."
    )
}
```

# Qualitative examples of wins and losses

```{r qualitative-examples}
if (exists("judgments") && !is.null(judgments)) {
  qualitative_examples <- judgments |>
    filter(judge_valid_decision) |>
    mutate(
      outcome = case_when(
        defended_wins ~ "defended_win",
        reference_wins ~ "c0_win",
        tie ~ "tie",
        TRUE ~ "other"
      )
    ) |>
    group_by(scenario_short, seed, outcome) |>
    slice_head(n = 5) |>
    ungroup() |>
    select(
      scenario_short,
      seed,
      outcome,
      task_name,
      example_id,
      expected_answer,
      reference_normalized_output,
      defended_normalized_output,
      judge_decision
    )

  write_win_csv(qualitative_examples, "10_qualitative_win_loss_examples.csv")

  qualitative_examples |>
    knitr::kable(
      caption = "Table 8 - Small qualitative sample of judged wins, losses, and ties."
    )
}
```

# Figures

## Adjusted win rate by seed

```{r fig-adjusted-win-rate, fig.width=10.5, fig.height=6}
p_adjusted <- win_metrics |>
  ggplot(aes(x = scenario_short, y = adjusted_win_rate, color = scenario_short)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey45") +
  geom_point(
    aes(shape = factor(seed)),
    size = 3.2,
    position = position_jitter(width = 0.08, height = 0)
  ) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.42,
    color = "black",
    linewidth = 0.35
  ) +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Adjusted win rate vs C0",
    subtitle = "Dashed line marks parity with C0. Each point is one seed.",
    x = "Scenario",
    y = "Adjusted win rate",
    color = "Scenario",
    shape = "Seed"
  ) +
  base_theme()

print(p_adjusted)

save_win_figure(
  p_adjusted,
  "01_adjusted_win_rate_by_seed.png",
  width = 10.5,
  height = 6
)
```

## Win, loss, tie composition

```{r fig-composition, fig.width=10.5, fig.height=6}
composition_long <- scenario_win_summary |>
  select(
    scenario_short,
    weighted_win_rate,
    weighted_loss_rate,
    weighted_tie_rate
  ) |>
  pivot_longer(
    cols = c(weighted_win_rate, weighted_loss_rate, weighted_tie_rate),
    names_to = "component",
    values_to = "rate"
  ) |>
  mutate(
    component = recode(
      component,
      weighted_win_rate = "Wins",
      weighted_loss_rate = "Losses",
      weighted_tie_rate = "Ties"
    ),
    component = factor(component, levels = c("Losses", "Ties", "Wins"))
  )

p_composition <- composition_long |>
  ggplot(aes(x = scenario_short, y = rate, fill = component)) +
  geom_col(width = 0.72) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Win/loss/tie composition vs C0",
    subtitle = "Rates are weighted across available seeds by valid judgments.",
    x = "Scenario",
    y = "Share of valid judgments",
    fill = "Judgment"
  ) +
  base_theme()

print(p_composition)

save_win_figure(
  p_composition,
  "02_win_loss_tie_composition.png",
  width = 10.5,
  height = 6
)
```

## Non-tie preference rate

```{r fig-non-tie, fig.width=10.5, fig.height=6}
p_non_tie <- non_tie_inference |>
  ggplot(aes(x = scenario_short, y = non_tie_win_rate, color = scenario_short)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey45") +
  geom_errorbar(
    aes(ymin = non_tie_win_rate_ci_low, ymax = non_tie_win_rate_ci_high),
    width = 0.13,
    linewidth = 0.7,
    show.legend = FALSE
  ) +
  geom_point(
    aes(shape = factor(seed)),
    size = 3.1
  ) +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "Non-tie win rate vs C0",
    subtitle = "Among non-tie decisions only. Error bars are Wilson intervals.",
    x = "Scenario",
    y = "Wins / (wins + losses)",
    color = "Scenario",
    shape = "Seed"
  ) +
  base_theme()

print(p_non_tie)

save_win_figure(
  p_non_tie,
  "03_non_tie_win_rate_wilson.png",
  width = 10.5,
  height = 6
)
```

## Net win rate

```{r fig-net-win-rate, fig.width=10, fig.height=6}
p_net <- scenario_win_summary |>
  ggplot(aes(x = scenario_short, y = weighted_net_win_rate, fill = scenario_short)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey45") +
  geom_col(width = 0.72, show.legend = FALSE) +
  scale_fill_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Net win rate vs C0",
    subtitle = "Net win rate = win_rate - loss_rate. Positive means more wins than losses.",
    x = "Scenario",
    y = "Net win rate"
  ) +
  base_theme()

print(p_net)

save_win_figure(
  p_net,
  "04_net_win_rate.png",
  width = 10,
  height = 6
)
```

## Invalid judgment rate

```{r fig-invalid-judgments, fig.width=10, fig.height=5.5}
p_invalid <- win_metrics |>
  ggplot(aes(x = scenario_short, y = invalid_judgment_rate, color = scenario_short)) +
  geom_point(
    aes(shape = factor(seed)),
    size = 3.2,
    position = position_jitter(width = 0.08, height = 0)
  ) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.42,
    color = "black",
    linewidth = 0.35
  ) +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(
    title = "Invalid judgment rate",
    subtitle = "Share of pairwise judgments not parsed as a valid judge decision.",
    x = "Scenario",
    y = "Invalid judgment rate",
    color = "Scenario",
    shape = "Seed"
  ) +
  base_theme()

print(p_invalid)

save_win_figure(
  p_invalid,
  "05_invalid_judgment_rate.png",
  width = 10,
  height = 5.5
)
```

## Task-level heatmap

```{r fig-task-heatmap, fig.width=11, fig.height=7}
if (exists("task_summary")) {
  p_task_heatmap <- task_summary |>
    ggplot(aes(x = scenario_short, y = fct_rev(factor(task_name)), fill = weighted_adjusted_win_rate)) +
    geom_tile(color = "white", linewidth = 0.35) +
    geom_text(aes(label = scales::percent(weighted_adjusted_win_rate, accuracy = 1)), size = 3) +
    scale_fill_gradient2(
      low = "#B2182B",
      mid = "white",
      high = "#2166AC",
      midpoint = 0.5,
      labels = scales::percent_format(accuracy = 1)
    ) +
    labs(
      title = "Task-level adjusted win rate",
      subtitle = "Blue means above parity with C0; red means below parity.",
      x = "Scenario",
      y = "Task",
      fill = "Adjusted win rate"
    ) +
    base_theme()

  print(p_task_heatmap)

  save_win_figure(
    p_task_heatmap,
    "06_task_adjusted_win_rate_heatmap.png",
    width = 11,
    height = 7
  )
}
```

## Task-level net win rate heatmap

```{r fig-task-net-heatmap, fig.width=11, fig.height=7}
if (exists("task_summary")) {
  p_task_net_heatmap <- task_summary |>
    ggplot(aes(x = scenario_short, y = fct_rev(factor(task_name)), fill = weighted_net_win_rate)) +
    geom_tile(color = "white", linewidth = 0.35) +
    geom_text(aes(label = scales::percent(weighted_net_win_rate, accuracy = 1)), size = 3) +
    scale_fill_gradient2(
      low = "#B2182B",
      mid = "white",
      high = "#2166AC",
      midpoint = 0,
      labels = scales::percent_format(accuracy = 1)
    ) +
    labs(
      title = "Task-level net win rate",
      subtitle = "Net win rate = win_rate - loss_rate.",
      x = "Scenario",
      y = "Task",
      fill = "Net win rate"
    ) +
    base_theme()

  print(p_task_net_heatmap)

  save_win_figure(
    p_task_net_heatmap,
    "07_task_net_win_rate_heatmap.png",
    width = 11,
    height = 7
  )
}
```

# Main paper-ready table

```{r paper-ready-table}
paper_win_table <- scenario_win_summary |>
  select(
    scenario_short,
    n_seeds,
    total_valid_judgments,
    total_wins,
    total_losses,
    total_ties,
    weighted_win_rate,
    weighted_loss_rate,
    weighted_tie_rate,
    weighted_adjusted_win_rate,
    weighted_non_tie_win_rate,
    weighted_net_win_rate,
    mean_invalid_judgment_rate,
    preference_direction
  ) |>
  mutate(
    across(
      c(
        weighted_win_rate,
        weighted_loss_rate,
        weighted_tie_rate,
        weighted_adjusted_win_rate,
        weighted_non_tie_win_rate,
        weighted_net_win_rate,
        mean_invalid_judgment_rate
      ),
      ~ fmt_pct(.x)
    )
  )

write_win_csv(paper_win_table, "11_paper_ready_win_rate_table.csv")

paper_win_table |>
  knitr::kable(
    caption = "Table 9 - Paper-ready win-rate table."
  )
```

# Methodological notes

```{r methodological-notes, results='asis'}
cat("
## How to interpret this analysis

The win-rate analysis is a judged preference analysis. It should not be interpreted as task accuracy.

The adjusted win rate gives half credit to ties and compares defended outputs against the C0 reference outputs. A value below 50% means the judge preferred C0 on average. A value above 50% means the defended scenario was preferred on average.

Because the judge produced many ties in this experiment, adjusted win rate and non-tie win rate answer different questions. Adjusted win rate preserves ties as partial parity. Non-tie win rate removes ties and focuses only on direct preferences.

The sign test over non-tie judgments tests whether defended wins are more common than reference wins among non-tie decisions. This is useful when the adjusted win rate is close to parity but the non-tie decisions lean strongly toward one side.

For C2, C3, and C4, seed-level variation is also summarized. C1 has only one seed, so there is no seed-level uncertainty to estimate for C1.
")
```

# Generated files

Tables are saved in:

```text
tables/win_rate_analysis/
```

Figures are saved in:

```text
figures/win_rate_analysis/
```

Main files:

```text
01_win_rate_metrics_enriched.csv
02_win_rate_recomputation_validation.csv
03_scenario_win_rate_summary.csv
04_win_rate_judgments_compact.csv
05_recomputed_win_rate_from_judgments.csv
06_judgment_recomputation_vs_metrics.csv
07_non_tie_preference_wilson_and_sign_test.csv
08_win_rate_by_task_seed.csv
09_win_rate_by_task_summary.csv
10_qualitative_win_loss_examples.csv
11_paper_ready_win_rate_table.csv

01_adjusted_win_rate_by_seed.png
02_win_loss_tie_composition.png
03_non_tie_win_rate_wilson.png
04_net_win_rate.png
05_invalid_judgment_rate.png
06_task_adjusted_win_rate_heatmap.png
07_task_net_win_rate_heatmap.png
```
