---
title: "RPC-PI Experiment - Accuracy and Task Utility Analysis"
author: ""
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    theme: sandstone
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)

options(knitr.kable.NA = "-")

required_packages <- c(
  "tidyverse",
  "readr",
  "ggplot2",
  "scales",
  "stringr",
  "forcats",
  "fs",
  "here",
  "knitr"
)

missing_packages <- required_packages[!required_packages %in% rownames(installed.packages())]
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

library(tidyverse)
library(readr)
library(ggplot2)
library(scales)
library(stringr)
library(forcats)
library(fs)
library(here)
library(knitr)
```

# Objective

This R Markdown focuses on **accuracy and task utility**.

It is different from ASV and Win Rate:

- ASV measures whether the attack reached the adversarial target;
- Win Rate measures judged preference against C0;
- Accuracy measures whether the model preserved the original task objective.

At the example level:

```text
task_success = normalized_output == expected_answer
```

In clean examples, this becomes:

```text
clean_accuracy = mean(task_success)
```

In attacked examples, this becomes:

```text
robust_accuracy = mean(task_success)
```

This report analyzes:

- clean accuracy;
- robust accuracy;
- task success rate across splits;
- seen vs unseen robust accuracy;
- accuracy by task;
- accuracy by attack type;
- utility drop vs C0;
- robust accuracy delta vs C0;
- paired clean improvements/regressions against C0;
- paired robust wins against C0;
- accuracy and ASV trade-off.

Higher accuracy is better.

# Paths and configuration

```{r paths}
DATA_DIR <- here::here("data", "results")
DESC_DIR <- here::here("tables", "descriptive")
ERROR_TABLE_DIR <- file.path(DATA_DIR, "error_analysis", "full", "tables")

OUT_TABLE_DIR <- here::here("tables", "accuracy_analysis")
OUT_FIG_DIR <- here::here("figures", "accuracy_analysis")

fs::dir_create(OUT_TABLE_DIR)
fs::dir_create(OUT_FIG_DIR)

if (!dir.exists(DATA_DIR)) {
  stop(
    "Could not find data/results/. Extract the ZIP first with: ",
    "unzip('results.zip', exdir = 'data')"
  )
}
```

# Helper functions

```{r helpers}
read_first_existing <- function(paths, required = TRUE) {
  existing <- paths[file.exists(paths)]

  if (length(existing) == 0) {
    if (required) {
      stop("None of these files were found:\n", paste(paths, collapse = "\n"))
    }
    return(NULL)
  }

  readr::read_csv(existing[[1]], show_col_types = FALSE)
}

coalesce_col <- function(df, cols, default = NA_real_) {
  existing <- cols[cols %in% names(df)]

  if (length(existing) == 0) {
    return(rep(default, nrow(df)))
  }

  out <- df[[existing[[1]]]]

  if (length(existing) > 1) {
    for (nm in existing[-1]) {
      out <- dplyr::coalesce(out, df[[nm]])
    }
  }

  out
}

scenario_order <- c(
  "c0_base",
  "c1_struq_format_only",
  "c2_struq_sft",
  "c3_secalign_dpo",
  "c4_ih_sft"
)

scenario_labels <- tibble(
  scenario_id = scenario_order,
  scenario_short = c("C0", "C1", "C2", "C3", "C4"),
  scenario_label = c(
    "C0 - Base model",
    "C1 - StruQ format-only",
    "C2 - StruQ-like SFT",
    "C3 - SecAlign-like DPO",
    "C4 - Instruction-Hierarchy-like SFT"
  )
)

scenario_palette <- c(
  C0 = "#7F7F7F",
  C1 = "#D55E00",
  C2 = "#0072B2",
  C3 = "#CC79A7",
  C4 = "#009E73"
)

sort_scenarios <- function(df) {
  df |>
    mutate(
      scenario_id = factor(as.character(scenario_id), levels = scenario_order),
      scenario_short = factor(as.character(scenario_short), levels = c("C0", "C1", "C2", "C3", "C4"))
    ) |>
    arrange(scenario_id)
}

add_scenario_info <- function(df) {
  out <- df |>
    mutate(scenario_id = as.character(scenario_id)) |>
    left_join(scenario_labels, by = "scenario_id", suffix = c("", "_lookup"))

  if (!("scenario_label" %in% names(out))) {
    out$scenario_label <- out$scenario_id
  }

  if ("scenario_label_lookup" %in% names(out)) {
    out <- out |>
      mutate(
        scenario_label = coalesce(
          as.character(scenario_label),
          as.character(scenario_label_lookup)
        )
      )
  }

  if (!("scenario_short" %in% names(out))) {
    out$scenario_short <- NA_character_
  }

  if ("scenario_short_lookup" %in% names(out)) {
    out <- out |>
      mutate(
        scenario_short = coalesce(
          as.character(scenario_short),
          as.character(scenario_short_lookup)
        )
      )
  }

  out |>
    mutate(
      scenario_short = ifelse(
        is.na(scenario_short),
        recode(
          as.character(scenario_id),
          c0_base = "C0",
          c1_struq_format_only = "C1",
          c2_struq_sft = "C2",
          c3_secalign_dpo = "C3",
          c4_ih_sft = "C4",
          .default = as.character(scenario_id)
        ),
        scenario_short
      )
    ) |>
    select(-any_of(c("scenario_label_lookup", "scenario_short_lookup"))) |>
    sort_scenarios()
}

to_logical_safe <- function(x) {
  if (is.logical(x)) return(x)

  x_chr <- tolower(as.character(x))

  case_when(
    x_chr %in% c("true", "t", "1", "yes") ~ TRUE,
    x_chr %in% c("false", "f", "0", "no") ~ FALSE,
    TRUE ~ NA
  )
}

safe_mean <- function(x) {
  if (sum(!is.na(x)) == 0) return(NA_real_)
  mean(x, na.rm = TRUE)
}

safe_sd <- function(x) {
  if (sum(!is.na(x)) <= 1) return(NA_real_)
  sd(x, na.rm = TRUE)
}

safe_sem <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) return(NA_real_)
  sd(x, na.rm = TRUE) / sqrt(n)
}

safe_tcrit <- function(n, confidence = 0.95) {
  if (is.na(n) || n <= 1) return(NA_real_)
  alpha <- 1 - confidence
  qt(1 - alpha / 2, df = n - 1)
}

wilson_ci <- function(k, n, conf.level = 0.95) {
  if (is.na(k) || is.na(n) || n <= 0) {
    return(tibble(estimate = NA_real_, ci_low = NA_real_, ci_high = NA_real_))
  }

  z <- qnorm(1 - (1 - conf.level) / 2)
  p <- k / n
  denom <- 1 + z^2 / n
  center <- (p + z^2 / (2 * n)) / denom
  half_width <- z * sqrt((p * (1 - p) / n) + (z^2 / (4 * n^2))) / denom

  tibble(
    estimate = p,
    ci_low = max(0, center - half_width),
    ci_high = min(1, center + half_width)
  )
}

safe_binom_p <- function(x, n, p = 0.5, alternative = "greater") {
  if (is.na(x) || is.na(n) || n <= 0 || x < 0 || x > n) {
    return(NA_real_)
  }

  tryCatch(
    binom.test(as.integer(x), as.integer(n), p = p, alternative = alternative)$p.value,
    error = function(e) NA_real_
  )
}

bootstrap_stratified_mean <- function(values, strata, B = 500) {
  values <- as.numeric(values)
  strata <- as.character(strata)

  valid <- !is.na(values) & !is.na(strata)

  values <- values[valid]
  strata <- strata[valid]

  if (length(values) == 0) {
    return(rep(NA_real_, B))
  }

  split_values <- split(values, strata)

  replicate(
    B,
    {
      sampled <- unlist(
        lapply(
          split_values,
          function(x) {
            sample(x, size = length(x), replace = TRUE)
          }
        ),
        use.names = FALSE
      )

      mean(sampled, na.rm = TRUE)
    }
  )
}

summarise_bootstrap <- function(values, strata, B = 500) {
  boot <- bootstrap_stratified_mean(values, strata, B = B)

  tibble(
    estimate = mean(values, na.rm = TRUE),
    ci_low = as.numeric(quantile(boot, 0.025, na.rm = TRUE)),
    ci_high = as.numeric(quantile(boot, 0.975, na.rm = TRUE))
  )
}

fmt_pct <- function(x, accuracy = 0.1) {
  scales::percent(x, accuracy = accuracy)
}

base_theme <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      legend.position = "bottom",
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(color = "grey30"),
      panel.grid.minor = element_blank(),
      axis.title = element_text(face = "bold")
    )
}

write_acc_csv <- function(df, filename) {
  readr::write_csv(df, file.path(OUT_TABLE_DIR, filename))
  invisible(df)
}

save_acc_figure <- function(plot, filename, width = 10, height = 6) {
  ggsave(
    file.path(OUT_FIG_DIR, filename),
    plot = plot,
    width = width,
    height = height,
    dpi = 320,
    bg = "white"
  )
  invisible(plot)
}
```

# Load inference examples

```{r load-inference}
inference <- read_first_existing(c(
  file.path(DESC_DIR, "01_inference_compact.csv")
)) |>
  add_scenario_info()

if (!("eval_group" %in% names(inference))) {
  inference$eval_group <- dplyr::case_when(
    inference$eval_split == "test_clean" ~ "clean",
    inference$eval_split == "test_attacked_seen" ~ "attacked_seen",
    inference$eval_split == "test_attacked_unseen" ~ "attacked_unseen",
    TRUE ~ as.character(inference$eval_split)
  )
}

if (!("attack_type" %in% names(inference))) {
  inference$attack_type <- NA_character_
}

if (!("is_correct_flag" %in% names(inference))) {
  if (all(c("normalized_output", "expected_answer") %in% names(inference))) {
    inference$is_correct_flag <- inference$normalized_output == inference$expected_answer
  } else {
    stop("Could not find is_correct_flag or normalized_output/expected_answer to compute accuracy.")
  }
}

if (!("followed_attack_flag" %in% names(inference))) {
  if (all(c("normalized_output", "attack_target") %in% names(inference))) {
    inference$followed_attack_flag <- inference$normalized_output == inference$attack_target
  } else {
    inference$followed_attack_flag <- NA
  }
}

if (!("invalid_output_flag" %in% names(inference))) {
  inference$invalid_output_flag <- NA
}

examples <- inference |>
  mutate(
    seed = as.integer(seed),
    eval_split = as.character(eval_split),
    eval_group = as.character(eval_group),
    split_group = case_when(
      eval_split == "test_clean" | eval_group == "clean" ~ "clean",
      eval_split == "test_attacked_seen" | eval_group == "attacked_seen" ~ "seen",
      eval_split == "test_attacked_unseen" | eval_group == "attacked_unseen" ~ "unseen",
      TRUE ~ as.character(eval_group)
    ),
    is_attacked_split = split_group %in% c("seen", "unseen"),
    task_name = as.character(task_name),
    attack_type = ifelse(is.na(attack_type) | attack_type == "", "clean", as.character(attack_type)),
    is_correct_flag = to_logical_safe(is_correct_flag),
    followed_attack_flag = to_logical_safe(followed_attack_flag),
    invalid_output_flag = to_logical_safe(invalid_output_flag),
    task_success = is_correct_flag,
    task_success_int = as.integer(task_success),
    asv = followed_attack_flag,
    asv_int = as.integer(asv),
    invalid_output_int = as.integer(invalid_output_flag)
  ) |>
  filter(!is.na(task_success))

write_acc_csv(
  examples |>
    select(
      scenario_id,
      scenario_short,
      seed,
      eval_split,
      split_group,
      example_id,
      task_name,
      attack_type,
      expected_answer,
      attack_target,
      normalized_output,
      task_success,
      asv,
      invalid_output_flag
    ),
  "01_examples_with_task_success.csv"
)

examples |>
  count(scenario_short, seed, split_group, name = "n_examples") |>
  knitr::kable(
    caption = "Table 1 - Examples used in accuracy analysis."
  )
```

# Accuracy by seed and split

```{r accuracy-by-seed-split}
accuracy_by_seed_split <- examples |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    seed,
    eval_split,
    split_group
  ) |>
  summarise(
    n = n(),
    correct = sum(task_success_int, na.rm = TRUE),
    task_success_rate = correct / n,
    asv_successes = sum(asv_int, na.rm = TRUE),
    asv_rate = if_else(is_attacked_split[1], asv_successes / n, NA_real_),
    invalid_outputs = sum(invalid_output_int, na.rm = TRUE),
    invalid_output_rate = safe_mean(invalid_output_int),
    .groups = "drop"
  ) |>
  rowwise() |>
  mutate(
    wilson = list(wilson_ci(correct, n, conf.level = 0.95))
  ) |>
  unnest(wilson) |>
  ungroup() |>
  rename(
    accuracy_wilson = estimate,
    accuracy_ci_low = ci_low,
    accuracy_ci_high = ci_high
  ) |>
  mutate(
    metric_name = case_when(
      split_group == "clean" ~ "clean_accuracy",
      split_group %in% c("seen", "unseen") ~ "robust_accuracy",
      TRUE ~ "task_success_rate"
    )
  ) |>
  sort_scenarios()

accuracy_all_attacks_seed <- examples |>
  filter(is_attacked_split) |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    seed
  ) |>
  summarise(
    eval_split = "all_attacks",
    split_group = "all_attacks",
    n = n(),
    correct = sum(task_success_int, na.rm = TRUE),
    task_success_rate = correct / n,
    asv_successes = sum(asv_int, na.rm = TRUE),
    asv_rate = asv_successes / n,
    invalid_outputs = sum(invalid_output_int, na.rm = TRUE),
    invalid_output_rate = safe_mean(invalid_output_int),
    .groups = "drop"
  ) |>
  rowwise() |>
  mutate(
    wilson = list(wilson_ci(correct, n, conf.level = 0.95))
  ) |>
  unnest(wilson) |>
  ungroup() |>
  rename(
    accuracy_wilson = estimate,
    accuracy_ci_low = ci_low,
    accuracy_ci_high = ci_high
  ) |>
  mutate(metric_name = "robust_accuracy_all_attacks") |>
  sort_scenarios()

accuracy_seed_summary <- bind_rows(accuracy_by_seed_split, accuracy_all_attacks_seed) |>
  mutate(
    split_group = factor(split_group, levels = c("clean", "seen", "unseen", "all_attacks"))
  ) |>
  sort_scenarios()

write_acc_csv(accuracy_seed_summary, "02_accuracy_by_seed_split_fraction.csv")

accuracy_seed_summary |>
  select(
    scenario_short,
    seed,
    split_group,
    n,
    correct,
    task_success_rate,
    accuracy_ci_low,
    accuracy_ci_high,
    asv_rate,
    invalid_output_rate
  ) |>
  mutate(
    across(
      c(task_success_rate, accuracy_ci_low, accuracy_ci_high, asv_rate, invalid_output_rate),
      ~ fmt_pct(.x)
    )
  ) |>
  knitr::kable(
    caption = "Table 2 - Accuracy by seed and split with Wilson intervals."
  )
```

# Scenario-level accuracy and utility summary

```{r scenario-summary}
accuracy_scenario_summary <- accuracy_seed_summary |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    split_group
  ) |>
  summarise(
    n_seeds = n_distinct(seed),
    seeds = paste(sort(unique(seed)), collapse = ", "),
    total_examples = sum(n, na.rm = TRUE),
    total_correct = sum(correct, na.rm = TRUE),
    weighted_accuracy = total_correct / total_examples,
    mean_accuracy = mean(task_success_rate, na.rm = TRUE),
    sd_accuracy = safe_sd(task_success_rate),
    sem_accuracy = safe_sem(task_success_rate),
    tcrit_95 = safe_tcrit(n_distinct(seed)),
    seed_ci_low = mean_accuracy - tcrit_95 * sem_accuracy,
    seed_ci_high = mean_accuracy + tcrit_95 * sem_accuracy,
    total_asv_successes = sum(asv_successes, na.rm = TRUE),
    weighted_asv_rate = if_else(
      first(as.character(split_group)) == "clean",
      NA_real_,
      total_asv_successes / total_examples
    ),
    weighted_invalid_output_rate = if_else(
      sum(n, na.rm = TRUE) > 0,
      sum(invalid_outputs, na.rm = TRUE) / sum(n, na.rm = TRUE),
      NA_real_
    ),
    .groups = "drop"
  ) |>
  rowwise() |>
  mutate(
    wilson = list(wilson_ci(total_correct, total_examples, conf.level = 0.95))
  ) |>
  unnest(wilson) |>
  ungroup() |>
  rename(
    weighted_accuracy_wilson = estimate,
    weighted_accuracy_ci_low = ci_low,
    weighted_accuracy_ci_high = ci_high
  ) |>
  group_by(split_group) |>
  mutate(
    c0_accuracy = weighted_accuracy[scenario_id == "c0_base"][1],
    accuracy_delta_vs_c0 = weighted_accuracy - c0_accuracy,
    utility_drop_vs_c0 = c0_accuracy - weighted_accuracy,
    relative_accuracy_vs_c0 = if_else(
      !is.na(c0_accuracy) & c0_accuracy > 0,
      weighted_accuracy / c0_accuracy,
      NA_real_
    )
  ) |>
  ungroup() |>
  sort_scenarios()

write_acc_csv(accuracy_scenario_summary, "03_accuracy_scenario_summary_fraction.csv")

accuracy_scenario_summary |>
  filter(split_group %in% c("clean", "all_attacks")) |>
  select(
    scenario_short,
    split_group,
    n_seeds,
    total_examples,
    total_correct,
    weighted_accuracy,
    weighted_accuracy_ci_low,
    weighted_accuracy_ci_high,
    accuracy_delta_vs_c0,
    utility_drop_vs_c0,
    relative_accuracy_vs_c0,
    weighted_asv_rate
  ) |>
  mutate(
    across(
      c(
        weighted_accuracy,
        weighted_accuracy_ci_low,
        weighted_accuracy_ci_high,
        accuracy_delta_vs_c0,
        utility_drop_vs_c0,
        relative_accuracy_vs_c0,
        weighted_asv_rate
      ),
      ~ fmt_pct(.x)
    )
  ) |>
  knitr::kable(
    caption = "Table 3 - Scenario-level clean and robust accuracy summary."
  )
```

# Clean utility vs robust utility

```{r clean-vs-robust-table}
clean_robust_summary <- accuracy_scenario_summary |>
  filter(split_group %in% c("clean", "all_attacks")) |>
  select(
    scenario_id,
    scenario_short,
    scenario_label,
    split_group,
    weighted_accuracy,
    weighted_accuracy_ci_low,
    weighted_accuracy_ci_high,
    accuracy_delta_vs_c0,
    utility_drop_vs_c0,
    relative_accuracy_vs_c0,
    weighted_asv_rate
  ) |>
  pivot_wider(
    names_from = split_group,
    values_from = c(
      weighted_accuracy,
      weighted_accuracy_ci_low,
      weighted_accuracy_ci_high,
      accuracy_delta_vs_c0,
      utility_drop_vs_c0,
      relative_accuracy_vs_c0,
      weighted_asv_rate
    )
  ) |>
  mutate(
    robust_minus_clean_accuracy = weighted_accuracy_all_attacks - weighted_accuracy_clean,
    robust_retention = if_else(
      !is.na(weighted_accuracy_clean) & weighted_accuracy_clean > 0,
      weighted_accuracy_all_attacks / weighted_accuracy_clean,
      NA_real_
    )
  ) |>
  sort_scenarios()

write_acc_csv(clean_robust_summary, "04_clean_vs_robust_utility_summary.csv")

clean_robust_summary |>
  select(
    scenario_short,
    weighted_accuracy_clean,
    accuracy_delta_vs_c0_clean,
    utility_drop_vs_c0_clean,
    weighted_accuracy_all_attacks,
    accuracy_delta_vs_c0_all_attacks,
    utility_drop_vs_c0_all_attacks,
    robust_minus_clean_accuracy,
    robust_retention,
    weighted_asv_rate_all_attacks
  ) |>
  mutate(across(where(is.numeric), ~ fmt_pct(.x))) |>
  knitr::kable(
    caption = "Table 4 - Clean utility vs robust utility."
  )
```

# Accuracy by task

```{r by-task}
accuracy_by_task <- examples |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    split_group,
    task_name
  ) |>
  summarise(
    n = n(),
    correct = sum(task_success_int, na.rm = TRUE),
    accuracy = correct / n,
    asv_rate = if_else(
      split_group[1] == "clean",
      NA_real_,
      sum(asv_int, na.rm = TRUE) / n
    ),
    .groups = "drop"
  ) |>
  rowwise() |>
  mutate(wilson = list(wilson_ci(correct, n, conf.level = 0.95))) |>
  unnest(wilson) |>
  ungroup() |>
  rename(
    accuracy_wilson = estimate,
    accuracy_ci_low = ci_low,
    accuracy_ci_high = ci_high
  ) |>
  group_by(split_group, task_name) |>
  mutate(
    c0_accuracy = accuracy[scenario_id == "c0_base"][1],
    accuracy_delta_vs_c0 = accuracy - c0_accuracy
  ) |>
  ungroup() |>
  sort_scenarios()

accuracy_by_task_all_attacks <- examples |>
  filter(is_attacked_split) |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    task_name
  ) |>
  summarise(
    split_group = "all_attacks",
    n = n(),
    correct = sum(task_success_int, na.rm = TRUE),
    accuracy = correct / n,
    asv_rate = sum(asv_int, na.rm = TRUE) / n,
    .groups = "drop"
  ) |>
  rowwise() |>
  mutate(wilson = list(wilson_ci(correct, n, conf.level = 0.95))) |>
  unnest(wilson) |>
  ungroup() |>
  rename(
    accuracy_wilson = estimate,
    accuracy_ci_low = ci_low,
    accuracy_ci_high = ci_high
  ) |>
  group_by(split_group, task_name) |>
  mutate(
    c0_accuracy = accuracy[scenario_id == "c0_base"][1],
    accuracy_delta_vs_c0 = accuracy - c0_accuracy
  ) |>
  ungroup() |>
  sort_scenarios()

accuracy_by_task_full <- bind_rows(accuracy_by_task, accuracy_by_task_all_attacks)

write_acc_csv(accuracy_by_task_full, "05_accuracy_by_task_fraction.csv")

accuracy_by_task_full |>
  filter(split_group %in% c("clean", "all_attacks")) |>
  select(
    scenario_short,
    split_group,
    task_name,
    n,
    correct,
    accuracy,
    accuracy_ci_low,
    accuracy_ci_high,
    accuracy_delta_vs_c0,
    asv_rate
  ) |>
  mutate(
    across(c(accuracy, accuracy_ci_low, accuracy_ci_high, accuracy_delta_vs_c0, asv_rate), ~ fmt_pct(.x))
  ) |>
  knitr::kable(
    caption = "Table 5 - Accuracy by task."
  )
```

# Accuracy by attack type

```{r by-attack-type}
accuracy_by_attack_type <- examples |>
  filter(is_attacked_split) |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    attack_type
  ) |>
  summarise(
    n = n(),
    correct = sum(task_success_int, na.rm = TRUE),
    robust_accuracy = correct / n,
    asv_rate = sum(asv_int, na.rm = TRUE) / n,
    invalid_output_rate = safe_mean(invalid_output_int),
    .groups = "drop"
  ) |>
  rowwise() |>
  mutate(wilson = list(wilson_ci(correct, n, conf.level = 0.95))) |>
  unnest(wilson) |>
  ungroup() |>
  rename(
    robust_accuracy_wilson = estimate,
    robust_accuracy_ci_low = ci_low,
    robust_accuracy_ci_high = ci_high
  ) |>
  group_by(attack_type) |>
  mutate(
    c0_robust_accuracy = robust_accuracy[scenario_id == "c0_base"][1],
    robust_accuracy_delta_vs_c0 = robust_accuracy - c0_robust_accuracy
  ) |>
  ungroup() |>
  sort_scenarios()

write_acc_csv(accuracy_by_attack_type, "06_accuracy_by_attack_type_fraction.csv")

accuracy_by_attack_type |>
  select(
    scenario_short,
    attack_type,
    n,
    correct,
    robust_accuracy,
    robust_accuracy_ci_low,
    robust_accuracy_ci_high,
    robust_accuracy_delta_vs_c0,
    asv_rate,
    invalid_output_rate
  ) |>
  mutate(
    across(
      c(
        robust_accuracy,
        robust_accuracy_ci_low,
        robust_accuracy_ci_high,
        robust_accuracy_delta_vs_c0,
        asv_rate,
        invalid_output_rate
      ),
      ~ fmt_pct(.x)
    )
  ) |>
  knitr::kable(
    caption = "Table 6 - Robust accuracy by attack type."
  )
```

# Paired accuracy comparison against C0

This section uses `paired_vs_c0_rows.csv`. It compares the defended scenario and C0 on the same example.

For accuracy, positive delta means improvement:

```text
accuracy delta vs C0 = defended_correct - C0_correct
```

```{r load-paired}
paired_rows_path <- file.path(ERROR_TABLE_DIR, "paired_vs_c0_rows.csv")

if (!file.exists(paired_rows_path)) {
  paired_accuracy <- NULL
  cat("paired_vs_c0_rows.csv not found. Paired accuracy sections will be skipped.")
} else {
  paired_rows <- readr::read_csv(
    paired_rows_path,
    show_col_types = FALSE,
    col_select = any_of(c(
      "scenario_id",
      "scenario_label",
      "seed",
      "eval_split",
      "example_id",
      "example_key",
      "base_id",
      "task_name",
      "attack_type",
      "is_correct",
      "reference_is_correct",
      "followed_attack",
      "reference_followed_attack"
    ))
  ) |>
    add_scenario_info() |>
    mutate(
      seed = as.integer(seed),
      eval_split = as.character(eval_split),
      task_name = as.character(task_name),
      attack_type = ifelse(is.na(attack_type) | attack_type == "", "clean", as.character(attack_type)),
      is_correct = to_logical_safe(is_correct),
      reference_is_correct = to_logical_safe(reference_is_correct),
      followed_attack = to_logical_safe(followed_attack),
      reference_followed_attack = to_logical_safe(reference_followed_attack),
      split_group = case_when(
        eval_split == "test_clean" ~ "clean",
        eval_split == "test_attacked_seen" ~ "seen",
        eval_split == "test_attacked_unseen" ~ "unseen",
        TRUE ~ eval_split
      ),
      is_attacked_split = split_group %in% c("seen", "unseen")
    ) |>
    filter(!is.na(is_correct), !is.na(reference_is_correct))

  paired_accuracy <- paired_rows |>
    mutate(
      defended_correct = as.integer(is_correct),
      c0_correct = as.integer(reference_is_correct),
      accuracy_delta_vs_c0 = defended_correct - c0_correct,
      defense_better = defended_correct == 1 & c0_correct == 0,
      defense_worse = defended_correct == 0 & c0_correct == 1,
      same_outcome = defended_correct == c0_correct,
      stratum = paste(seed, eval_split, task_name, attack_type, sep = "__")
    )

  write_acc_csv(paired_accuracy, "07_paired_accuracy_rows_vs_c0.csv")

  paired_accuracy |>
    count(scenario_short, seed, split_group, name = "n_pairs") |>
    knitr::kable(
      caption = "Table 7 - Paired accuracy rows against C0."
    )
}
```

```{r paired-summary}
if (exists("paired_accuracy") && !is.null(paired_accuracy)) {
  paired_accuracy_summary <- paired_accuracy |>
    group_by(
      scenario_id,
      scenario_short,
      scenario_label,
      eval_split,
      split_group
    ) |>
    summarise(
      n_pairs = n(),
      defended_correct = sum(defended_correct, na.rm = TRUE),
      c0_correct = sum(c0_correct, na.rm = TRUE),
      defended_accuracy = defended_correct / n_pairs,
      c0_accuracy = c0_correct / n_pairs,
      accuracy_delta_vs_c0 = defended_accuracy - c0_accuracy,
      better = sum(defense_better, na.rm = TRUE),
      worse = sum(defense_worse, na.rm = TRUE),
      same = sum(same_outcome, na.rm = TRUE),
      discordant_pairs = better + worse,
      better_rate_among_discordant = if_else(discordant_pairs > 0, better / discordant_pairs, NA_real_),
      sign_test_p_value = purrr::map2_dbl(
        better,
        discordant_pairs,
        ~ safe_binom_p(.x, .y, p = 0.5, alternative = "greater")
      ),
      significant_improvement_05 = !is.na(sign_test_p_value) & sign_test_p_value < 0.05,
      .groups = "drop"
    ) |>
    sort_scenarios()

  paired_accuracy_all_attacks <- paired_accuracy |>
    filter(is_attacked_split) |>
    group_by(
      scenario_id,
      scenario_short,
      scenario_label
    ) |>
    summarise(
      eval_split = "all_attacks",
      split_group = "all_attacks",
      n_pairs = n(),
      defended_correct = sum(defended_correct, na.rm = TRUE),
      c0_correct = sum(c0_correct, na.rm = TRUE),
      defended_accuracy = defended_correct / n_pairs,
      c0_accuracy = c0_correct / n_pairs,
      accuracy_delta_vs_c0 = defended_accuracy - c0_accuracy,
      better = sum(defense_better, na.rm = TRUE),
      worse = sum(defense_worse, na.rm = TRUE),
      same = sum(same_outcome, na.rm = TRUE),
      discordant_pairs = better + worse,
      better_rate_among_discordant = if_else(discordant_pairs > 0, better / discordant_pairs, NA_real_),
      sign_test_p_value = purrr::map2_dbl(
        better,
        discordant_pairs,
        ~ safe_binom_p(.x, .y, p = 0.5, alternative = "greater")
      ),
      significant_improvement_05 = !is.na(sign_test_p_value) & sign_test_p_value < 0.05,
      .groups = "drop"
    ) |>
    sort_scenarios()

  paired_accuracy_summary_full <- bind_rows(paired_accuracy_summary, paired_accuracy_all_attacks) |>
    mutate(
      split_group = factor(split_group, levels = c("clean", "seen", "unseen", "all_attacks"))
    ) |>
    sort_scenarios()

  write_acc_csv(paired_accuracy_summary_full, "08_paired_accuracy_summary_vs_c0.csv")

  paired_accuracy_summary_full |>
    filter(split_group %in% c("clean", "all_attacks")) |>
    select(
      scenario_short,
      split_group,
      n_pairs,
      defended_accuracy,
      c0_accuracy,
      accuracy_delta_vs_c0,
      better,
      worse,
      same,
      better_rate_among_discordant,
      sign_test_p_value,
      significant_improvement_05
    ) |>
    mutate(
      across(
        c(defended_accuracy, c0_accuracy, accuracy_delta_vs_c0, better_rate_among_discordant),
        ~ fmt_pct(.x)
      ),
      sign_test_p_value = scales::pvalue(sign_test_p_value, accuracy = 0.001)
    ) |>
    knitr::kable(
      caption = "Table 8 - Paired accuracy summary against C0."
    )
}
```

# Bootstrap paired accuracy delta

```{r paired-bootstrap}
N_BOOTSTRAP <- 500
set.seed(42)

if (exists("paired_accuracy") && !is.null(paired_accuracy)) {
  paired_bootstrap_accuracy <- paired_accuracy |>
    group_by(
      scenario_id,
      scenario_short,
      scenario_label,
      eval_split,
      split_group
    ) |>
    group_modify(
      ~ summarise_bootstrap(
        values = .x$accuracy_delta_vs_c0,
        strata = .x$stratum,
        B = N_BOOTSTRAP
      )
    ) |>
    ungroup() |>
    rename(
      mean_accuracy_delta_vs_c0 = estimate,
      boot_ci_low = ci_low,
      boot_ci_high = ci_high
    ) |>
    sort_scenarios()

  paired_bootstrap_accuracy_all <- paired_accuracy |>
    filter(is_attacked_split) |>
    group_by(
      scenario_id,
      scenario_short,
      scenario_label
    ) |>
    group_modify(
      ~ summarise_bootstrap(
        values = .x$accuracy_delta_vs_c0,
        strata = .x$stratum,
        B = N_BOOTSTRAP
      )
    ) |>
    ungroup() |>
    mutate(
      eval_split = "all_attacks",
      split_group = "all_attacks"
    ) |>
    rename(
      mean_accuracy_delta_vs_c0 = estimate,
      boot_ci_low = ci_low,
      boot_ci_high = ci_high
    ) |>
    sort_scenarios()

  paired_bootstrap_accuracy_full <- bind_rows(paired_bootstrap_accuracy, paired_bootstrap_accuracy_all) |>
    mutate(
      split_group = factor(split_group, levels = c("clean", "seen", "unseen", "all_attacks"))
    )

  write_acc_csv(paired_bootstrap_accuracy_full, "09_paired_bootstrap_accuracy_delta_vs_c0.csv")

  paired_bootstrap_accuracy_full |>
    filter(split_group %in% c("clean", "all_attacks")) |>
    select(
      scenario_short,
      split_group,
      mean_accuracy_delta_vs_c0,
      boot_ci_low,
      boot_ci_high
    ) |>
    mutate(
      across(c(mean_accuracy_delta_vs_c0, boot_ci_low, boot_ci_high), ~ fmt_pct(.x))
    ) |>
    knitr::kable(
      caption = "Table 9 - Bootstrap CI for paired accuracy delta against C0."
    )
}
```

# Qualitative accuracy samples

```{r qualitative-samples}
accuracy_samples <- examples |>
  mutate(
    accuracy_category = case_when(
      split_group == "clean" & task_success ~ "clean_correct",
      split_group == "clean" & !task_success ~ "clean_wrong",
      is_attacked_split & task_success & !asv ~ "robust_correct_attack_blocked",
      is_attacked_split & !task_success & asv ~ "attack_followed_task_wrong",
      is_attacked_split & !task_success & !asv ~ "attack_blocked_task_wrong",
      TRUE ~ "other"
    )
  ) |>
  group_by(scenario_short, split_group, accuracy_category) |>
  slice_head(n = 5) |>
  ungroup() |>
  select(
    scenario_short,
    seed,
    split_group,
    accuracy_category,
    task_name,
    attack_type,
    example_id,
    expected_answer,
    attack_target,
    normalized_output,
    task_success,
    asv
  )

write_acc_csv(accuracy_samples, "10_qualitative_accuracy_samples.csv")

accuracy_samples |>
  knitr::kable(
    caption = "Table 10 - Small qualitative sample of accuracy outcomes."
  )
```

# Figures

## Clean and robust accuracy

```{r fig-clean-robust, fig.width=10.5, fig.height=6}
clean_robust_plot_data <- accuracy_scenario_summary |>
  filter(split_group %in% c("clean", "all_attacks")) |>
  mutate(
    split_label = recode(
      as.character(split_group),
      clean = "Clean accuracy",
      all_attacks = "Robust accuracy"
    )
  )

p_clean_robust <- clean_robust_plot_data |>
  ggplot(aes(x = scenario_short, y = weighted_accuracy, color = scenario_short)) +
  geom_errorbar(
    aes(ymin = weighted_accuracy_ci_low, ymax = weighted_accuracy_ci_high),
    width = 0.13,
    linewidth = 0.7,
    show.legend = FALSE
  ) +
  geom_point(size = 3.4, show.legend = FALSE) +
  facet_wrap(~ split_label, scales = "free_y") +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Clean and robust accuracy by scenario",
    subtitle = "Wilson intervals are computed over individual examples.",
    x = "Scenario",
    y = "Accuracy"
  ) +
  base_theme()

print(p_clean_robust)

save_acc_figure(
  p_clean_robust,
  "01_clean_and_robust_accuracy_wilson.png",
  width = 10.5,
  height = 6
)
```

## Seen vs unseen robust accuracy

```{r fig-seen-unseen, fig.width=10.5, fig.height=6}
p_seen_unseen <- accuracy_scenario_summary |>
  filter(split_group %in% c("seen", "unseen")) |>
  ggplot(aes(x = split_group, y = weighted_accuracy, color = scenario_short, group = scenario_short)) +
  geom_line(alpha = 0.65) +
  geom_errorbar(
    aes(ymin = weighted_accuracy_ci_low, ymax = weighted_accuracy_ci_high),
    width = 0.08,
    linewidth = 0.65
  ) +
  geom_point(size = 3) +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Robust accuracy on seen vs unseen attacks",
    subtitle = "Higher values indicate better preservation of the original task.",
    x = NULL,
    y = "Robust accuracy",
    color = "Scenario"
  ) +
  base_theme()

print(p_seen_unseen)

save_acc_figure(
  p_seen_unseen,
  "02_robust_accuracy_seen_vs_unseen.png",
  width = 10.5,
  height = 6
)
```

## Accuracy delta vs C0

```{r fig-delta-vs-c0, fig.width=10.5, fig.height=6}
p_delta <- accuracy_scenario_summary |>
  filter(split_group %in% c("clean", "all_attacks")) |>
  filter(scenario_id != "c0_base") |>
  mutate(
    split_label = recode(
      as.character(split_group),
      clean = "Clean accuracy delta",
      all_attacks = "Robust accuracy delta"
    )
  ) |>
  ggplot(aes(x = scenario_short, y = accuracy_delta_vs_c0, fill = scenario_short)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey45") +
  geom_col(width = 0.72, show.legend = FALSE) +
  facet_wrap(~ split_label, scales = "free_y") +
  scale_fill_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Accuracy delta vs C0",
    subtitle = "Positive values mean higher accuracy than C0.",
    x = "Scenario",
    y = "Scenario accuracy - C0 accuracy"
  ) +
  base_theme()

print(p_delta)

save_acc_figure(
  p_delta,
  "03_accuracy_delta_vs_c0.png",
  width = 10.5,
  height = 6
)
```

## Utility drop vs C0

```{r fig-utility-drop, fig.width=10.5, fig.height=6}
p_utility_drop <- accuracy_scenario_summary |>
  filter(split_group %in% c("clean", "all_attacks")) |>
  filter(scenario_id != "c0_base") |>
  mutate(
    split_label = recode(
      as.character(split_group),
      clean = "Clean utility drop",
      all_attacks = "Robust utility drop"
    )
  ) |>
  ggplot(aes(x = scenario_short, y = utility_drop_vs_c0, fill = scenario_short)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey45") +
  geom_col(width = 0.72, show.legend = FALSE) +
  facet_wrap(~ split_label, scales = "free_y") +
  scale_fill_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Utility drop vs C0",
    subtitle = "Negative values mean the scenario is more accurate than C0.",
    x = "Scenario",
    y = "C0 accuracy - scenario accuracy"
  ) +
  base_theme()

print(p_utility_drop)

save_acc_figure(
  p_utility_drop,
  "04_utility_drop_vs_c0.png",
  width = 10.5,
  height = 6
)
```

## Accuracy by task heatmap

```{r fig-task-heatmap, fig.width=11, fig.height=7}
p_task_heatmap <- accuracy_by_task_full |>
  filter(split_group == "all_attacks") |>
  ggplot(aes(x = scenario_short, y = fct_rev(factor(task_name)), fill = accuracy)) +
  geom_tile(color = "white", linewidth = 0.35) +
  geom_text(aes(label = scales::percent(accuracy, accuracy = 1)), size = 3) +
  scale_fill_gradient2(
    low = "#B2182B",
    mid = "white",
    high = "#2166AC",
    midpoint = 0.5,
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Robust accuracy by task",
    subtitle = "Blue means higher accuracy; red means lower accuracy.",
    x = "Scenario",
    y = "Task",
    fill = "Accuracy"
  ) +
  base_theme()

print(p_task_heatmap)

save_acc_figure(
  p_task_heatmap,
  "05_robust_accuracy_by_task_heatmap.png",
  width = 11,
  height = 7
)
```

## Accuracy by attack type heatmap

```{r fig-attack-type-heatmap, fig.width=11, fig.height=7}
p_attack_heatmap <- accuracy_by_attack_type |>
  ggplot(aes(x = scenario_short, y = fct_rev(factor(attack_type)), fill = robust_accuracy)) +
  geom_tile(color = "white", linewidth = 0.35) +
  geom_text(aes(label = scales::percent(robust_accuracy, accuracy = 1)), size = 3) +
  scale_fill_gradient2(
    low = "#B2182B",
    mid = "white",
    high = "#2166AC",
    midpoint = 0.5,
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Robust accuracy by attack type",
    subtitle = "Blue means higher accuracy; red means lower accuracy.",
    x = "Scenario",
    y = "Attack type",
    fill = "Robust accuracy"
  ) +
  base_theme()

print(p_attack_heatmap)

save_acc_figure(
  p_attack_heatmap,
  "06_robust_accuracy_by_attack_type_heatmap.png",
  width = 11,
  height = 7
)
```

## Clean accuracy vs robust accuracy

```{r fig-clean-vs-robust-scatter, fig.width=9, fig.height=6.5}
p_clean_vs_robust <- clean_robust_summary |>
  ggplot(aes(x = weighted_accuracy_clean, y = weighted_accuracy_all_attacks, color = scenario_short, label = scenario_short)) +
  geom_point(size = 4) +
  geom_text(nudge_y = 0.025, show.legend = FALSE) +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Clean accuracy vs robust accuracy",
    subtitle = "Ideal region is upper-right: high clean utility and high robustness.",
    x = "Clean accuracy",
    y = "Robust accuracy",
    color = "Scenario"
  ) +
  base_theme()

print(p_clean_vs_robust)

save_acc_figure(
  p_clean_vs_robust,
  "07_clean_accuracy_vs_robust_accuracy.png",
  width = 9,
  height = 6.5
)
```

## Robust accuracy vs ASV

```{r fig-accuracy-vs-asv, fig.width=9, fig.height=6.5}
p_accuracy_asv <- clean_robust_summary |>
  ggplot(aes(x = weighted_asv_rate_all_attacks, y = weighted_accuracy_all_attacks, color = scenario_short, label = scenario_short)) +
  geom_point(size = 4) +
  geom_text(nudge_y = 0.025, show.legend = FALSE) +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Robust accuracy vs ASV",
    subtitle = "Ideal region is upper-left: high task utility and low attack success.",
    x = "ASV",
    y = "Robust accuracy",
    color = "Scenario"
  ) +
  base_theme()

print(p_accuracy_asv)

save_acc_figure(
  p_accuracy_asv,
  "08_robust_accuracy_vs_asv.png",
  width = 9,
  height = 6.5
)
```

## Paired accuracy delta bootstrap

```{r fig-paired-bootstrap, fig.width=10.5, fig.height=6}
if (exists("paired_bootstrap_accuracy_full")) {
  p_paired_boot <- paired_bootstrap_accuracy_full |>
    filter(split_group %in% c("clean", "all_attacks")) |>
    mutate(
      split_label = recode(
        as.character(split_group),
        clean = "Clean",
        all_attacks = "All attacks"
      )
    ) |>
    ggplot(aes(x = scenario_short, y = mean_accuracy_delta_vs_c0, color = scenario_short)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey45") +
    geom_errorbar(
      aes(ymin = boot_ci_low, ymax = boot_ci_high),
      width = 0.13,
      linewidth = 0.7,
      show.legend = FALSE
    ) +
    geom_point(size = 3.3, show.legend = FALSE) +
    facet_wrap(~ split_label, scales = "free_y") +
    scale_color_manual(values = scenario_palette, drop = FALSE) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(
      title = "Paired accuracy delta vs C0",
      subtitle = "Bootstrap intervals. Positive values mean higher accuracy than C0.",
      x = "Scenario",
      y = "Accuracy delta vs C0"
    ) +
    base_theme()

  print(p_paired_boot)

  save_acc_figure(
    p_paired_boot,
    "09_paired_accuracy_delta_bootstrap.png",
    width = 10.5,
    height = 6
  )
}
```

## Discordant-pair accuracy sign test

```{r fig-discordant-sign-test, fig.width=10.5, fig.height=6}
if (exists("paired_accuracy_summary_full")) {
  p_discordant <- paired_accuracy_summary_full |>
    filter(split_group %in% c("clean", "all_attacks")) |>
    mutate(
      split_label = recode(
        as.character(split_group),
        clean = "Clean",
        all_attacks = "All attacks"
      )
    ) |>
    ggplot(aes(x = scenario_short, y = better_rate_among_discordant, fill = scenario_short)) +
    geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey45") +
    geom_col(width = 0.72, show.legend = FALSE) +
    facet_wrap(~ split_label) +
    scale_fill_manual(values = scenario_palette, drop = FALSE) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1)) +
    labs(
      title = "Accuracy improvement among discordant pairs",
      subtitle = "A pair improves when the defended scenario is correct and C0 is wrong.",
      x = "Scenario",
      y = "Better among discordant pairs"
    ) +
    base_theme()

  print(p_discordant)

  save_acc_figure(
    p_discordant,
    "10_accuracy_discordant_pair_sign_test.png",
    width = 10.5,
    height = 6
  )
}
```

# Paper-ready table

```{r paper-ready-table}
paper_accuracy_table <- clean_robust_summary |>
  select(
    scenario_short,
    weighted_accuracy_clean,
    weighted_accuracy_ci_low_clean,
    weighted_accuracy_ci_high_clean,
    accuracy_delta_vs_c0_clean,
    utility_drop_vs_c0_clean,
    weighted_accuracy_all_attacks,
    weighted_accuracy_ci_low_all_attacks,
    weighted_accuracy_ci_high_all_attacks,
    accuracy_delta_vs_c0_all_attacks,
    utility_drop_vs_c0_all_attacks,
    robust_retention,
    weighted_asv_rate_all_attacks
  ) |>
  mutate(
    across(where(is.numeric), ~ fmt_pct(.x))
  )

write_acc_csv(paper_accuracy_table, "11_paper_ready_accuracy_table.csv")

paper_accuracy_table |>
  knitr::kable(
    caption = "Table 11 - Paper-ready accuracy and task utility table."
  )
```

# Methodological notes

```{r methodological-notes, results='asis'}
cat("
## How to interpret accuracy and utility

Clean accuracy measures normal task utility without an attack. Robust accuracy measures task utility under prompt injection.

Task success is computed as normalized_output == expected_answer. Therefore, robust accuracy is not just the absence of attack following. A model can block the attack and still fail the original task; that would reduce ASV but not count as robustly correct.

Utility drop vs C0 is computed as C0 accuracy - scenario accuracy. Negative utility drop means the scenario is more accurate than C0.

The Wilson intervals in this report are example-level intervals for proportions. The paired sections compare each defended scenario against C0 on the same examples.
")
```

# Generated files

Tables are saved in:

```text
tables/accuracy_analysis/
```

Figures are saved in:

```text
figures/accuracy_analysis/
```

Main files:

```text
01_examples_with_task_success.csv
02_accuracy_by_seed_split_fraction.csv
03_accuracy_scenario_summary_fraction.csv
04_clean_vs_robust_utility_summary.csv
05_accuracy_by_task_fraction.csv
06_accuracy_by_attack_type_fraction.csv
07_paired_accuracy_rows_vs_c0.csv
08_paired_accuracy_summary_vs_c0.csv
09_paired_bootstrap_accuracy_delta_vs_c0.csv
10_qualitative_accuracy_samples.csv
11_paper_ready_accuracy_table.csv

01_clean_and_robust_accuracy_wilson.png
02_robust_accuracy_seen_vs_unseen.png
03_accuracy_delta_vs_c0.png
04_utility_drop_vs_c0.png
05_robust_accuracy_by_task_heatmap.png
06_robust_accuracy_by_attack_type_heatmap.png
07_clean_accuracy_vs_robust_accuracy.png
08_robust_accuracy_vs_asv.png
09_paired_accuracy_delta_bootstrap.png
10_accuracy_discordant_pair_sign_test.png
```
