---
title: "RPC-PI Experiment - Inference Time Analysis"
author: ""
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    theme: sandstone
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)

options(knitr.kable.NA = "-")

required_packages <- c(
  "tidyverse",
  "readr",
  "ggplot2",
  "scales",
  "stringr",
  "forcats",
  "fs",
  "here",
  "knitr",
  "jsonlite",
  "lubridate"
)

missing_packages <- required_packages[!required_packages %in% rownames(installed.packages())]
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

library(tidyverse)
library(readr)
library(ggplot2)
library(scales)
library(stringr)
library(forcats)
library(fs)
library(here)
library(knitr)
library(jsonlite)
library(lubridate)
```

# Objective

This R Markdown analyzes **inference cost**.

The main questions are:

```text
How long does inference take for each scenario?
Does the defense add inference overhead?
Which split is more expensive?
How many examples per second does each scenario process?
How does inference cost relate to task utility and attack success?
```

This report uses:

```text
logs/inference/05_run_inference_summary.json
logs/inference/05_run_inference_events.jsonl
logs/inference/runs/.../run_metadata.json
logs/inference/runs/.../test_*_metadata.json
```

It also optionally joins automatic metrics from:

```text
tables/descriptive/01_inference_compact.csv
```

when that file exists.

# Expected folder structure

This Rmd expects the logs extracted into one of these locations:

```text
data/logs/inference/
logs/inference/
```

If `logs.zip` exists in the project root or inside `data/`, this Rmd tries to extract it automatically to `data/`.

# Paths

```{r paths}
OUT_TABLE_DIR <- here::here("tables", "inference_time_analysis")
OUT_FIG_DIR <- here::here("figures", "inference_time_analysis")

fs::dir_create(OUT_TABLE_DIR)
fs::dir_create(OUT_FIG_DIR)

# Try to extract logs.zip automatically if needed.
if (!dir.exists(here::here("data", "logs")) && file.exists(here::here("logs.zip"))) {
  dir.create(here::here("data"), showWarnings = FALSE)
  unzip(here::here("logs.zip"), exdir = here::here("data"))
}

if (!dir.exists(here::here("data", "logs")) && file.exists(here::here("data", "logs.zip"))) {
  dir.create(here::here("data"), showWarnings = FALSE)
  unzip(here::here("data", "logs.zip"), exdir = here::here("data"))
}

candidate_inference_dirs <- c(
  here::here("data", "logs", "inference"),
  here::here("logs", "inference")
)

INFERENCE_LOG_DIR <- candidate_inference_dirs[dir.exists(candidate_inference_dirs)][1]

if (is.na(INFERENCE_LOG_DIR)) {
  stop(
    "Could not find inference logs. Expected one of:\n",
    paste(candidate_inference_dirs, collapse = "\n"),
    "\nOr place logs.zip in the project root or data/."
  )
}

SUMMARY_PATH <- file.path(INFERENCE_LOG_DIR, "05_run_inference_summary.json")
EVENTS_PATH <- file.path(INFERENCE_LOG_DIR, "05_run_inference_events.jsonl")
RUNS_DIR <- file.path(INFERENCE_LOG_DIR, "runs")
DESC_INFERENCE_PATH <- here::here("tables", "descriptive", "01_inference_compact.csv")

if (!file.exists(SUMMARY_PATH)) {
  stop("Could not find inference summary JSON: ", SUMMARY_PATH)
}
```

# Helper functions

```{r helpers}
`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0) y else x
}

scenario_order <- c(
  "c0_base",
  "c1_struq_format_only",
  "c2_struq_sft",
  "c3_secalign_dpo",
  "c4_ih_sft"
)

scenario_labels <- tibble(
  scenario_id = scenario_order,
  scenario_short = c("C0", "C1", "C2", "C3", "C4"),
  scenario_label = c(
    "C0 - Base model",
    "C1 - StruQ format-only",
    "C2 - StruQ-like SFT",
    "C3 - SecAlign-like DPO",
    "C4 - Instruction-Hierarchy-like SFT"
  )
)

scenario_palette <- c(
  C0 = "#7F7F7F",
  C1 = "#D55E00",
  C2 = "#0072B2",
  C3 = "#CC79A7",
  C4 = "#009E73"
)

split_labels <- c(
  test_clean = "Clean",
  test_attacked_seen = "Attacked seen",
  test_attacked_unseen = "Attacked unseen",
  all_splits = "All splits",
  all_attacks = "All attacks"
)

split_order <- c(
  "test_clean",
  "test_attacked_seen",
  "test_attacked_unseen",
  "all_attacks",
  "all_splits"
)

sort_scenarios <- function(df) {
  df |>
    mutate(
      scenario_id = factor(as.character(scenario_id), levels = scenario_order),
      scenario_short = factor(as.character(scenario_short), levels = c("C0", "C1", "C2", "C3", "C4"))
    ) |>
    arrange(scenario_id)
}

add_scenario_info <- function(df) {
  out <- df |>
    mutate(scenario_id = as.character(scenario_id)) |>
    left_join(scenario_labels, by = "scenario_id", suffix = c("", "_lookup"))

  if (!("scenario_label" %in% names(out))) {
    out$scenario_label <- out$scenario_id
  }

  if ("scenario_label_lookup" %in% names(out)) {
    out <- out |>
      mutate(
        scenario_label = coalesce(
          as.character(scenario_label),
          as.character(scenario_label_lookup)
        )
      )
  }

  if (!("scenario_short" %in% names(out))) {
    out$scenario_short <- NA_character_
  }

  if ("scenario_short_lookup" %in% names(out)) {
    out <- out |>
      mutate(
        scenario_short = coalesce(
          as.character(scenario_short),
          as.character(scenario_short_lookup)
        )
      )
  }

  out |>
    mutate(
      scenario_short = ifelse(
        is.na(scenario_short),
        recode(
          as.character(scenario_id),
          c0_base = "C0",
          c1_struq_format_only = "C1",
          c2_struq_sft = "C2",
          c3_secalign_dpo = "C3",
          c4_ih_sft = "C4",
          .default = as.character(scenario_id)
        ),
        scenario_short
      )
    ) |>
    select(-any_of(c("scenario_label_lookup", "scenario_short_lookup"))) |>
    sort_scenarios()
}

get_scalar <- function(obj, field, default = NA) {
  value <- obj[[field]]

  if (is.null(value) || length(value) == 0) {
    return(default)
  }

  if (is.data.frame(value)) {
    if (nrow(value) == 0 || ncol(value) == 0) return(default)
    return(value[[1]][[1]])
  }

  if (is.list(value) && length(value) == 1 && !is.null(value[[1]])) {
    return(value[[1]])
  }

  value[[1]]
}

get_nested_scalar <- function(obj, parent, child, default = NA) {
  flat_name <- paste0(parent, ".", child)

  flat_value <- obj[[flat_name]]

  if (!is.null(flat_value) && length(flat_value) > 0) {
    return(flat_value[[1]])
  }

  parent_value <- obj[[parent]]

  if (!is.null(parent_value)) {
    if (is.data.frame(parent_value) && child %in% names(parent_value)) {
      return(parent_value[[child]][[1]])
    }

    if (is.list(parent_value) && !is.null(parent_value[[child]])) {
      return(parent_value[[child]][[1]])
    }
  }

  default
}

replace_null_columns <- function(x) {
  if (!is.list(x)) {
    return(x)
  }

  n <- 1
  non_null_lengths <- purrr::map_int(
    x,
    function(value) {
      if (is.null(value)) {
        0L
      } else if (is.data.frame(value)) {
        nrow(value)
      } else {
        length(value)
      }
    }
  )

  if (any(non_null_lengths > 0)) {
    n <- max(non_null_lengths)
  }

  purrr::map(
    x,
    function(value) {
      if (is.null(value)) {
        rep(NA_character_, n)
      } else {
        value
      }
    }
  )
}

as_tibble_json_safe <- function(x) {
  if (is.null(x)) {
    return(tibble())
  }

  if (is.data.frame(x)) {
    x <- replace_null_columns(x)
    return(tibble::as_tibble(x))
  }

  if (is.list(x)) {
    x <- replace_null_columns(x)
    return(tibble::as_tibble(x))
  }

  tibble(value = x)
}

parse_time_utc <- function(x) {
  lubridate::ymd_hms(x, tz = "UTC", quiet = TRUE)
}

safe_mean <- function(x) {
  if (sum(!is.na(x)) == 0) return(NA_real_)
  mean(x, na.rm = TRUE)
}

safe_weighted_mean <- function(x, w) {
  x <- as.numeric(x)
  w <- as.numeric(w)

  ok <- !is.na(x) & !is.na(w) & w > 0

  if (sum(ok) == 0) {
    return(NA_real_)
  }

  weighted.mean(x[ok], w = w[ok], na.rm = TRUE)
}

safe_sd <- function(x) {
  if (sum(!is.na(x)) <= 1) return(NA_real_)
  sd(x, na.rm = TRUE)
}

safe_sem <- function(x) {
  n <- sum(!is.na(x))
  if (n <= 1) return(NA_real_)
  sd(x, na.rm = TRUE) / sqrt(n)
}

safe_tcrit <- function(n, confidence = 0.95) {
  if (is.na(n) || n <= 1) return(NA_real_)
  alpha <- 1 - confidence
  qt(1 - alpha / 2, df = n - 1)
}

to_logical_safe <- function(x) {
  if (is.logical(x)) return(x)

  x_chr <- tolower(as.character(x))

  case_when(
    x_chr %in% c("true", "t", "1", "yes") ~ TRUE,
    x_chr %in% c("false", "f", "0", "no") ~ FALSE,
    TRUE ~ NA
  )
}

fmt_duration <- function(seconds) {
  ifelse(
    is.na(seconds),
    NA_character_,
    paste0(
      floor(seconds / 3600), "h ",
      floor((seconds %% 3600) / 60), "m ",
      round(seconds %% 60), "s"
    )
  )
}

fmt_number <- function(x, accuracy = 0.01) {
  scales::number(x, accuracy = accuracy, big.mark = ",")
}

fmt_pct <- function(x, accuracy = 0.1) {
  scales::percent(x, accuracy = accuracy)
}

base_theme <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      legend.position = "bottom",
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(color = "grey30"),
      panel.grid.minor = element_blank(),
      axis.title = element_text(face = "bold")
    )
}

write_inference_csv <- function(df, filename) {
  readr::write_csv(df, file.path(OUT_TABLE_DIR, filename))
  invisible(df)
}

save_inference_figure <- function(plot, filename, width = 10, height = 6) {
  ggsave(
    file.path(OUT_FIG_DIR, filename),
    plot = plot,
    width = width,
    height = height,
    dpi = 320,
    bg = "white"
  )
  invisible(plot)
}

read_jsonl_robust <- function(path, bad_lines_filename = "00_bad_jsonl_lines.csv") {
  if (!file.exists(path)) {
    return(tibble())
  }

  lines <- readLines(path, warn = FALSE, encoding = "UTF-8")

  if (length(lines) == 0) {
    return(tibble())
  }

  parsed <- vector("list", length(lines))
  bad_lines <- list()

  for (i in seq_along(lines)) {
    line <- lines[[i]]

    if (is.na(line) || trimws(line) == "") {
      next
    }

    obj <- tryCatch(
      jsonlite::fromJSON(line, flatten = TRUE),
      error = function(e) {
        bad_lines[[length(bad_lines) + 1]] <<- tibble(
          line_number = i,
          error = conditionMessage(e),
          line_prefix = substr(line, 1, 250)
        )
        NULL
      }
    )

    if (!is.null(obj)) {
      parsed[[i]] <- as_tibble_json_safe(obj)
    }
  }

  parsed <- parsed[!vapply(parsed, is.null, logical(1))]
  bad_lines_df <- dplyr::bind_rows(bad_lines)

  if (nrow(bad_lines_df) > 0) {
    readr::write_csv(
      bad_lines_df,
      file.path(OUT_TABLE_DIR, bad_lines_filename)
    )
  }

  if (length(parsed) == 0) {
    return(tibble())
  }

  dplyr::bind_rows(parsed)
}

read_split_metadata_file <- function(path) {
  obj <- jsonlite::fromJSON(path, flatten = TRUE)

  tibble(
    scenario_id = as.character(get_scalar(obj, "scenario_id", NA_character_)),
    scenario_label_raw = as.character(get_scalar(obj, "scenario_label", NA_character_)),
    seed = as.integer(get_scalar(obj, "seed", NA_integer_)),
    split = as.character(get_scalar(obj, "split", NA_character_)),
    status = as.character(get_scalar(obj, "status", NA_character_)),
    run_mode = as.character(get_scalar(obj, "run_mode", NA_character_)),
    started_at_utc = as.character(get_scalar(obj, "started_at_utc", NA_character_)),
    finished_at_utc = as.character(get_scalar(obj, "finished_at_utc", NA_character_)),
    elapsed_seconds = as.numeric(get_scalar(obj, "elapsed_seconds", NA_real_)),
    rows = as.numeric(get_scalar(obj, "rows", NA_real_)),
    output_path = as.character(get_scalar(obj, "output_path", NA_character_)),
    adapter_path = as.character(get_scalar(obj, "adapter_path", NA_character_)),
    prompt_strategy = as.character(get_scalar(obj, "prompt_strategy", NA_character_)),
    normalized_null_count = as.numeric(get_scalar(obj, "normalized_null_count", NA_real_)),
    max_new_tokens = as.numeric(get_nested_scalar(obj, "generation_config", "max_new_tokens", NA_real_)),
    do_sample = as.character(get_nested_scalar(obj, "generation_config", "do_sample", NA_character_)),
    num_beams = as.numeric(get_nested_scalar(obj, "generation_config", "num_beams", NA_real_)),
    metadata_path = as.character(path)
  )
}

read_run_metadata_file <- function(path) {
  obj <- jsonlite::fromJSON(path, flatten = TRUE)

  tibble(
    scenario_id = as.character(get_scalar(obj, "scenario_id", NA_character_)),
    scenario_label_raw = as.character(get_scalar(obj, "scenario_label", NA_character_)),
    seed = as.integer(get_scalar(obj, "seed", NA_integer_)),
    status = as.character(get_scalar(obj, "status", NA_character_)),
    run_mode = as.character(get_scalar(obj, "run_mode", NA_character_)),
    started_at_utc = as.character(get_scalar(obj, "started_at_utc", NA_character_)),
    finished_at_utc = as.character(get_scalar(obj, "finished_at_utc", NA_character_)),
    elapsed_seconds = as.numeric(get_scalar(obj, "elapsed_seconds", NA_real_)),
    adapter_path = as.character(get_scalar(obj, "adapter_path", NA_character_)),
    prompt_strategy = as.character(get_scalar(obj, "prompt_strategy", NA_character_)),
    metadata_path = as.character(path)
  )
}
```

# Load inference summary

```{r load-summary}
inference_summary_json <- jsonlite::fromJSON(SUMMARY_PATH, flatten = TRUE)

generation_config <- inference_summary_json$generation_config %||% list()

experiment_metadata <- tibble(
  notebook = inference_summary_json$notebook %||% NA_character_,
  created_at_utc = inference_summary_json$created_at_utc %||% NA_character_,
  run_mode = inference_summary_json$run_mode %||% NA_character_,
  base_model_id = inference_summary_json$base_model_id %||% NA_character_,
  dataset_seed = inference_summary_json$dataset_seed %||% NA_integer_,
  max_input_length = inference_summary_json$max_input_length %||% NA_integer_,
  generation_batch_size = inference_summary_json$generation_batch_size %||% NA_integer_,
  use_4bit = inference_summary_json$use_4bit %||% NA,
  max_new_tokens = generation_config$max_new_tokens %||% NA_integer_,
  do_sample = generation_config$do_sample %||% NA,
  num_beams = generation_config$num_beams %||% NA_integer_,
  results_dir = inference_summary_json$results_dir %||% NA_character_,
  events_log_path = inference_summary_json$events_log_path %||% NA_character_,
  runs_log_dir = inference_summary_json$runs_log_dir %||% NA_character_
)

result_files <- as_tibble_json_safe(inference_summary_json$result_files) |>
  add_scenario_info() |>
  mutate(
    seed = as.integer(seed),
    split = as.character(split),
    rows = as.numeric(rows),
    exists = to_logical_safe(exists)
  ) |>
  sort_scenarios()

write_inference_csv(experiment_metadata, "01_inference_experiment_metadata.csv")
write_inference_csv(result_files, "02_inference_result_files_from_summary.csv")

experiment_metadata |>
  knitr::kable(
    caption = "Table 1 - Inference experiment metadata."
  )

result_files |>
  select(scenario_short, seed, split, rows, exists, path) |>
  knitr::kable(
    caption = "Table 2 - Result files listed in inference summary."
  )
```

# Load split metadata files

```{r load-split-metadata}
split_metadata_files <- if (dir.exists(RUNS_DIR)) {
  fs::dir_ls(RUNS_DIR, recurse = TRUE, regexp = "test_.*_metadata\\.json$")
} else {
  character()
}

split_metadata <- purrr::map_dfr(
  split_metadata_files,
  read_split_metadata_file
) |>
  add_scenario_info() |>
  mutate(
    split = as.character(split),
    split_label = recode(split, !!!split_labels, .default = split),
    split_label = factor(split_label, levels = split_labels[c("test_clean", "test_attacked_seen", "test_attacked_unseen")]),
    started_at = parse_time_utc(started_at_utc),
    finished_at = parse_time_utc(finished_at_utc),
    elapsed_minutes = elapsed_seconds / 60,
    seconds_per_example = elapsed_seconds / rows,
    examples_per_second = rows / elapsed_seconds,
    generated_token_budget = rows * max_new_tokens,
    token_budget_per_second = generated_token_budget / elapsed_seconds,
    normalized_null_rate = normalized_null_count / rows,
    is_attacked_split = split %in% c("test_attacked_seen", "test_attacked_unseen"),
    adapter_used = !is.na(adapter_path) & adapter_path != "",
    prompt_strategy = ifelse(is.na(prompt_strategy) | prompt_strategy == "", "unknown", prompt_strategy)
  ) |>
  sort_scenarios()

if (nrow(split_metadata) == 0) {
  stop("No test_*_metadata.json files were found under: ", RUNS_DIR)
}

write_inference_csv(split_metadata, "03_inference_split_metadata_enriched.csv")

split_metadata |>
  select(
    scenario_short,
    seed,
    split,
    status,
    rows,
    elapsed_minutes,
    seconds_per_example,
    examples_per_second,
    normalized_null_count,
    normalized_null_rate,
    prompt_strategy,
    adapter_used
  ) |>
  mutate(
    elapsed_minutes = fmt_number(elapsed_minutes),
    seconds_per_example = fmt_number(seconds_per_example, accuracy = 0.0001),
    examples_per_second = fmt_number(examples_per_second, accuracy = 0.001),
    normalized_null_rate = fmt_pct(normalized_null_rate)
  ) |>
  knitr::kable(
    caption = "Table 3 - Inference split metadata."
  )
```

# Load run metadata files

```{r load-run-metadata}
run_metadata_files <- if (dir.exists(RUNS_DIR)) {
  fs::dir_ls(RUNS_DIR, recurse = TRUE, regexp = "run_metadata\\.json$")
} else {
  character()
}

run_metadata <- purrr::map_dfr(
  run_metadata_files,
  read_run_metadata_file
) |>
  add_scenario_info() |>
  mutate(
    started_at = parse_time_utc(started_at_utc),
    finished_at = parse_time_utc(finished_at_utc),
    elapsed_minutes = elapsed_seconds / 60,
    adapter_used = !is.na(adapter_path) & adapter_path != "",
    prompt_strategy = ifelse(is.na(prompt_strategy) | prompt_strategy == "", "unknown", prompt_strategy)
  ) |>
  left_join(
    split_metadata |>
      group_by(scenario_id, seed) |>
      summarise(
        split_elapsed_seconds_sum = sum(elapsed_seconds, na.rm = TRUE),
        split_rows_sum = sum(rows, na.rm = TRUE),
        split_count = n(),
        .groups = "drop"
      ),
    by = c("scenario_id", "seed")
  ) |>
  mutate(
    run_overhead_seconds = elapsed_seconds - split_elapsed_seconds_sum,
    run_overhead_rate = run_overhead_seconds / elapsed_seconds,
    run_examples_per_second = split_rows_sum / elapsed_seconds,
    run_seconds_per_example = elapsed_seconds / split_rows_sum
  ) |>
  sort_scenarios()

write_inference_csv(run_metadata, "04_inference_run_metadata_enriched.csv")

run_metadata |>
  select(
    scenario_short,
    seed,
    status,
    elapsed_minutes,
    split_elapsed_seconds_sum,
    run_overhead_seconds,
    run_overhead_rate,
    split_rows_sum,
    run_examples_per_second,
    run_seconds_per_example,
    prompt_strategy,
    adapter_used
  ) |>
  mutate(
    elapsed_minutes = fmt_number(elapsed_minutes),
    split_elapsed_seconds_sum = fmt_number(split_elapsed_seconds_sum),
    run_overhead_seconds = fmt_number(run_overhead_seconds),
    run_overhead_rate = fmt_pct(run_overhead_rate),
    run_examples_per_second = fmt_number(run_examples_per_second, accuracy = 0.001),
    run_seconds_per_example = fmt_number(run_seconds_per_example, accuracy = 0.0001)
  ) |>
  knitr::kable(
    caption = "Table 4 - Inference run-level metadata."
  )
```

# Split-level inference cost summary

```{r split-summary}
split_summary_by_scenario <- split_metadata |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    split,
    split_label,
    prompt_strategy
  ) |>
  summarise(
    n_runs = n(),
    n_completed = sum(status == "completed", na.rm = TRUE),
    seeds = paste(sort(unique(seed)), collapse = ", "),
    rows_per_run = first(rows),
    total_rows = sum(rows, na.rm = TRUE),
    total_elapsed_seconds = sum(elapsed_seconds, na.rm = TRUE),
    mean_elapsed_seconds = mean(elapsed_seconds, na.rm = TRUE),
    sd_elapsed_seconds = safe_sd(elapsed_seconds),
    sem_elapsed_seconds = safe_sem(elapsed_seconds),
    tcrit_95 = safe_tcrit(n()),
    ci95_low_elapsed_seconds = mean_elapsed_seconds - tcrit_95 * sem_elapsed_seconds,
    ci95_high_elapsed_seconds = mean_elapsed_seconds + tcrit_95 * sem_elapsed_seconds,
    mean_seconds_per_example = mean(seconds_per_example, na.rm = TRUE),
    sd_seconds_per_example = safe_sd(seconds_per_example),
    mean_examples_per_second = mean(examples_per_second, na.rm = TRUE),
    sd_examples_per_second = safe_sd(examples_per_second),
    mean_token_budget_per_second = mean(token_budget_per_second, na.rm = TRUE),
    total_normalized_null_count = sum(normalized_null_count, na.rm = TRUE),
    normalized_null_rate = total_normalized_null_count / total_rows,
    .groups = "drop"
  ) |>
  mutate(
    mean_elapsed_minutes = mean_elapsed_seconds / 60,
    total_elapsed_minutes = total_elapsed_seconds / 60,
    ci95_low_elapsed_minutes = ci95_low_elapsed_seconds / 60,
    ci95_high_elapsed_minutes = ci95_high_elapsed_seconds / 60,
    mean_duration_label = fmt_duration(mean_elapsed_seconds),
    total_duration_label = fmt_duration(total_elapsed_seconds)
  ) |>
  sort_scenarios()

write_inference_csv(split_summary_by_scenario, "05_split_summary_by_scenario.csv")

split_summary_by_scenario |>
  select(
    scenario_short,
    split,
    prompt_strategy,
    n_runs,
    seeds,
    rows_per_run,
    mean_elapsed_minutes,
    mean_seconds_per_example,
    mean_examples_per_second,
    mean_token_budget_per_second,
    normalized_null_rate,
    mean_duration_label
  ) |>
  mutate(
    mean_elapsed_minutes = fmt_number(mean_elapsed_minutes),
    mean_seconds_per_example = fmt_number(mean_seconds_per_example, accuracy = 0.0001),
    mean_examples_per_second = fmt_number(mean_examples_per_second, accuracy = 0.001),
    mean_token_budget_per_second = fmt_number(mean_token_budget_per_second, accuracy = 0.001),
    normalized_null_rate = fmt_pct(normalized_null_rate)
  ) |>
  knitr::kable(
    caption = "Table 5 - Split-level inference cost by scenario."
  )
```

# Scenario-level inference cost summary

```{r scenario-summary}
scenario_all_splits <- split_metadata |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    seed,
    prompt_strategy
  ) |>
  summarise(
    split = "all_splits",
    split_label = "All splits",
    rows = sum(rows, na.rm = TRUE),
    elapsed_seconds = sum(elapsed_seconds, na.rm = TRUE),
    normalized_null_count = sum(normalized_null_count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    seconds_per_example = elapsed_seconds / rows,
    examples_per_second = rows / elapsed_seconds,
    normalized_null_rate = normalized_null_count / rows
  )

scenario_all_attacks <- split_metadata |>
  filter(is_attacked_split) |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    seed,
    prompt_strategy
  ) |>
  summarise(
    split = "all_attacks",
    split_label = "All attacks",
    rows = sum(rows, na.rm = TRUE),
    elapsed_seconds = sum(elapsed_seconds, na.rm = TRUE),
    normalized_null_count = sum(normalized_null_count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    seconds_per_example = elapsed_seconds / rows,
    examples_per_second = rows / elapsed_seconds,
    normalized_null_rate = normalized_null_count / rows
  )

aggregate_split_metadata <- bind_rows(
  split_metadata |>
    select(
      scenario_id,
      scenario_short,
      scenario_label,
      seed,
      prompt_strategy,
      split,
      split_label,
      rows,
      elapsed_seconds,
      seconds_per_example,
      examples_per_second,
      normalized_null_count,
      normalized_null_rate
    ),
  scenario_all_attacks,
  scenario_all_splits
) |>
  mutate(
    split = factor(as.character(split), levels = split_order),
    split_label = factor(as.character(split_label), levels = split_labels[split_order])
  ) |>
  sort_scenarios()

scenario_inference_summary <- aggregate_split_metadata |>
  group_by(
    scenario_id,
    scenario_short,
    scenario_label,
    split,
    split_label,
    prompt_strategy
  ) |>
  summarise(
    n_runs = n(),
    seeds = paste(sort(unique(seed)), collapse = ", "),
    total_rows = sum(rows, na.rm = TRUE),
    mean_rows = mean(rows, na.rm = TRUE),
    total_elapsed_seconds = sum(elapsed_seconds, na.rm = TRUE),
    mean_elapsed_seconds = mean(elapsed_seconds, na.rm = TRUE),
    sd_elapsed_seconds = safe_sd(elapsed_seconds),
    mean_seconds_per_example = mean(seconds_per_example, na.rm = TRUE),
    sd_seconds_per_example = safe_sd(seconds_per_example),
    mean_examples_per_second = mean(examples_per_second, na.rm = TRUE),
    sd_examples_per_second = safe_sd(examples_per_second),
    total_normalized_null_count = sum(normalized_null_count, na.rm = TRUE),
    normalized_null_rate = total_normalized_null_count / total_rows,
    .groups = "drop"
  ) |>
  mutate(
    mean_elapsed_minutes = mean_elapsed_seconds / 60,
    total_elapsed_minutes = total_elapsed_seconds / 60,
    total_elapsed_hours = total_elapsed_seconds / 3600,
    mean_duration_label = fmt_duration(mean_elapsed_seconds),
    total_duration_label = fmt_duration(total_elapsed_seconds)
  ) |>
  sort_scenarios()

write_inference_csv(aggregate_split_metadata, "06_aggregate_split_metadata.csv")
write_inference_csv(scenario_inference_summary, "07_scenario_inference_summary.csv")

scenario_inference_summary |>
  filter(split %in% c("all_attacks", "all_splits")) |>
  select(
    scenario_short,
    split_label,
    prompt_strategy,
    n_runs,
    seeds,
    mean_rows,
    mean_elapsed_minutes,
    mean_seconds_per_example,
    mean_examples_per_second,
    normalized_null_rate,
    mean_duration_label
  ) |>
  mutate(
    mean_elapsed_minutes = fmt_number(mean_elapsed_minutes),
    mean_seconds_per_example = fmt_number(mean_seconds_per_example, accuracy = 0.0001),
    mean_examples_per_second = fmt_number(mean_examples_per_second, accuracy = 0.001),
    normalized_null_rate = fmt_pct(normalized_null_rate)
  ) |>
  knitr::kable(
    caption = "Table 6 - Scenario-level inference cost summary."
  )
```

# Inference overhead against C0

```{r overhead-vs-c0}
c0_split_reference <- scenario_inference_summary |>
  filter(scenario_id == "c0_base") |>
  select(
    split,
    c0_mean_seconds_per_example = mean_seconds_per_example,
    c0_mean_examples_per_second = mean_examples_per_second,
    c0_mean_elapsed_seconds = mean_elapsed_seconds
  )

overhead_vs_c0 <- scenario_inference_summary |>
  left_join(c0_split_reference, by = "split") |>
  mutate(
    seconds_per_example_delta_vs_c0 = mean_seconds_per_example - c0_mean_seconds_per_example,
    relative_seconds_per_example_vs_c0 = mean_seconds_per_example / c0_mean_seconds_per_example,
    overhead_rate_vs_c0 = relative_seconds_per_example_vs_c0 - 1,
    examples_per_second_delta_vs_c0 = mean_examples_per_second - c0_mean_examples_per_second,
    relative_examples_per_second_vs_c0 = mean_examples_per_second / c0_mean_examples_per_second,
    elapsed_time_delta_vs_c0 = mean_elapsed_seconds - c0_mean_elapsed_seconds,
    relative_elapsed_time_vs_c0 = mean_elapsed_seconds / c0_mean_elapsed_seconds
  ) |>
  sort_scenarios()

write_inference_csv(overhead_vs_c0, "08_inference_overhead_vs_c0.csv")

overhead_vs_c0 |>
  filter(split %in% c("test_clean", "test_attacked_seen", "test_attacked_unseen", "all_splits")) |>
  select(
    scenario_short,
    split_label,
    mean_seconds_per_example,
    c0_mean_seconds_per_example,
    seconds_per_example_delta_vs_c0,
    relative_seconds_per_example_vs_c0,
    overhead_rate_vs_c0,
    mean_examples_per_second,
    relative_examples_per_second_vs_c0
  ) |>
  mutate(
    mean_seconds_per_example = fmt_number(mean_seconds_per_example, accuracy = 0.0001),
    c0_mean_seconds_per_example = fmt_number(c0_mean_seconds_per_example, accuracy = 0.0001),
    seconds_per_example_delta_vs_c0 = fmt_number(seconds_per_example_delta_vs_c0, accuracy = 0.0001),
    relative_seconds_per_example_vs_c0 = fmt_number(relative_seconds_per_example_vs_c0, accuracy = 0.001),
    overhead_rate_vs_c0 = fmt_pct(overhead_rate_vs_c0),
    mean_examples_per_second = fmt_number(mean_examples_per_second, accuracy = 0.001),
    relative_examples_per_second_vs_c0 = fmt_number(relative_examples_per_second_vs_c0, accuracy = 0.001)
  ) |>
  knitr::kable(
    caption = "Table 7 - Inference overhead against C0."
  )
```

# Prompt strategy and adapter overhead

```{r prompt-strategy-summary}
prompt_strategy_summary <- aggregate_split_metadata |>
  group_by(prompt_strategy, split, split_label) |>
  summarise(
    n_runs = n(),
    scenarios = paste(sort(unique(as.character(scenario_short))), collapse = ", "),
    total_rows = sum(rows, na.rm = TRUE),
    total_elapsed_seconds = sum(elapsed_seconds, na.rm = TRUE),
    mean_seconds_per_example = mean(seconds_per_example, na.rm = TRUE),
    mean_examples_per_second = mean(examples_per_second, na.rm = TRUE),
    normalized_null_rate = sum(normalized_null_count, na.rm = TRUE) / total_rows,
    .groups = "drop"
  ) |>
  mutate(
    total_elapsed_minutes = total_elapsed_seconds / 60
  )

write_inference_csv(prompt_strategy_summary, "09_prompt_strategy_inference_summary.csv")

prompt_strategy_summary |>
  filter(split %in% c("test_clean", "all_attacks", "all_splits")) |>
  select(
    prompt_strategy,
    split_label,
    scenarios,
    n_runs,
    total_rows,
    total_elapsed_minutes,
    mean_seconds_per_example,
    mean_examples_per_second,
    normalized_null_rate
  ) |>
  mutate(
    total_elapsed_minutes = fmt_number(total_elapsed_minutes),
    mean_seconds_per_example = fmt_number(mean_seconds_per_example, accuracy = 0.0001),
    mean_examples_per_second = fmt_number(mean_examples_per_second, accuracy = 0.001),
    normalized_null_rate = fmt_pct(normalized_null_rate)
  ) |>
  knitr::kable(
    caption = "Table 8 - Prompt strategy inference summary."
  )
```

# Events and errors

```{r events-and-errors}
inference_events <- read_jsonl_robust(
  EVENTS_PATH,
  bad_lines_filename = "00_bad_inference_event_lines.csv"
)

if (nrow(inference_events) > 0) {
  if (!("scenario_id" %in% names(inference_events)) && "scenario" %in% names(inference_events)) {
    inference_events <- inference_events |>
      rename(scenario_id = scenario)
  }

  inference_events <- inference_events |>
    add_scenario_info() |>
    mutate(
      seed = as.integer(seed),
      event_type = as.character(event_type),
      run_mode = as.character(run_mode),
      split = if ("split" %in% names(inference_events)) as.character(split) else NA_character_,
      timestamp = parse_time_utc(timestamp_utc),
      rows = if ("rows" %in% names(inference_events)) as.numeric(rows) else NA_real_,
      elapsed_seconds = if ("elapsed_seconds" %in% names(inference_events)) as.numeric(elapsed_seconds) else NA_real_,
      normalized_null_count = if ("normalized_null_count" %in% names(inference_events)) as.numeric(normalized_null_count) else NA_real_,
      examples_per_second = rows / elapsed_seconds,
      seconds_per_example = elapsed_seconds / rows
    ) |>
    sort_scenarios()

  event_counts <- inference_events |>
    count(run_mode, scenario_id, scenario_short, seed, event_type, name = "n_events") |>
    sort_scenarios()

  split_completion_events <- inference_events |>
    filter(event_type == "split_completed") |>
    mutate(
      normalized_null_rate = normalized_null_count / rows
    ) |>
    select(
      run_mode,
      scenario_id,
      scenario_short,
      seed,
      split,
      rows,
      elapsed_seconds,
      examples_per_second,
      seconds_per_example,
      normalized_null_count,
      normalized_null_rate,
      timestamp
    ) |>
    sort_scenarios()

  write_inference_csv(inference_events, "10_inference_events_enriched.csv")
  write_inference_csv(event_counts, "11_inference_event_counts.csv")
  write_inference_csv(split_completion_events, "12_split_completion_events.csv")

  event_counts |>
    knitr::kable(
      caption = "Table 9 - Inference event counts."
    )
} else {
  cat("No inference events found.")
}

error_files <- if (dir.exists(RUNS_DIR)) {
  fs::dir_ls(RUNS_DIR, recurse = TRUE, regexp = "error\\.txt$")
} else {
  character()
}

inference_errors <- purrr::map_dfr(
  error_files,
  function(path) {
    parts <- stringr::str_split(as.character(path), "[/\\\\]", simplify = TRUE)
    runs_idx <- which(parts == "runs")

    scenario <- if (length(runs_idx) > 0) parts[runs_idx + 1] else NA_character_
    seed_part <- if (length(runs_idx) > 0) parts[runs_idx + 2] else NA_character_
    seed <- stringr::str_remove(seed_part, "^seed_")

    error_text <- paste(readLines(path, warn = FALSE, encoding = "UTF-8"), collapse = "\n")

    tibble(
      scenario_id = scenario,
      seed = as.integer(seed),
      error_path = as.character(path),
      error_text = error_text,
      error_preview = substr(error_text, 1, 500)
    )
  }
)

if (nrow(inference_errors) > 0) {
  inference_errors <- inference_errors |>
    add_scenario_info() |>
    sort_scenarios()

  write_inference_csv(inference_errors, "13_inference_error_files.csv")

  inference_errors |>
    select(scenario_short, seed, error_path, error_preview) |>
    knitr::kable(
      caption = "Table 10 - Error files found in inference run directories."
    )
} else {
  cat("No error.txt files found in inference run directories.")
}
```

# Optional join with quality metrics

```{r join-quality-metrics}
if (file.exists(DESC_INFERENCE_PATH)) {
  inference_compact <- readr::read_csv(DESC_INFERENCE_PATH, show_col_types = FALSE) |>
    add_scenario_info()

  if (!("is_correct_flag" %in% names(inference_compact))) {
    if (all(c("normalized_output", "expected_answer") %in% names(inference_compact))) {
      inference_compact$is_correct_flag <- inference_compact$normalized_output == inference_compact$expected_answer
    } else {
      inference_compact$is_correct_flag <- NA
    }
  }

  if (!("followed_attack_flag" %in% names(inference_compact))) {
    if (all(c("normalized_output", "attack_target") %in% names(inference_compact))) {
      inference_compact$followed_attack_flag <- inference_compact$normalized_output == inference_compact$attack_target
    } else {
      inference_compact$followed_attack_flag <- NA
    }
  }

  if (!("invalid_output_flag" %in% names(inference_compact))) {
    inference_compact$invalid_output_flag <- NA
  }

  quality_by_split <- inference_compact |>
    mutate(
      seed = as.integer(seed),
      split = as.character(eval_split),
      is_correct_flag = to_logical_safe(is_correct_flag),
      followed_attack_flag = to_logical_safe(followed_attack_flag),
      invalid_output_flag = to_logical_safe(invalid_output_flag),
      is_attacked_split = split %in% c("test_attacked_seen", "test_attacked_unseen")
    ) |>
    group_by(scenario_id, scenario_short, scenario_label, seed, split) |>
    summarise(
      metric_rows = n(),
      accuracy = mean(as.integer(is_correct_flag), na.rm = TRUE),
      asv_rate = if_else(
        first(is_attacked_split),
        mean(as.integer(followed_attack_flag), na.rm = TRUE),
        NA_real_
      ),
      invalid_output_rate = mean(as.integer(invalid_output_flag), na.rm = TRUE),
      .groups = "drop"
    ) |>
    sort_scenarios()

  quality_all_attacks <- quality_by_split |>
    filter(split %in% c("test_attacked_seen", "test_attacked_unseen")) |>
    group_by(scenario_id, scenario_short, scenario_label, seed) |>
    summarise(
      split = "all_attacks",
      accuracy = safe_weighted_mean(accuracy, metric_rows),
      asv_rate = safe_weighted_mean(asv_rate, metric_rows),
      invalid_output_rate = safe_weighted_mean(invalid_output_rate, metric_rows),
      metric_rows = sum(metric_rows, na.rm = TRUE),
      .groups = "drop"
    )

  quality_all_splits <- quality_by_split |>
    group_by(scenario_id, scenario_short, scenario_label, seed) |>
    summarise(
      split = "all_splits",
      accuracy = safe_weighted_mean(accuracy, metric_rows),
      asv_rate = safe_weighted_mean(asv_rate, metric_rows),
      invalid_output_rate = safe_weighted_mean(invalid_output_rate, metric_rows),
      metric_rows = sum(metric_rows, na.rm = TRUE),
      .groups = "drop"
    )

  quality_metrics_full <- bind_rows(
    quality_by_split,
    quality_all_attacks,
    quality_all_splits
  )

  inference_cost_quality <- aggregate_split_metadata |>
    left_join(
      quality_metrics_full |>
        select(
          scenario_id,
          seed,
          split,
          metric_rows,
          accuracy,
          asv_rate,
          invalid_output_rate
        ),
      by = c("scenario_id", "seed", "split")
    ) |>
    mutate(
      correct_examples_per_second = accuracy * examples_per_second,
      attack_successes_per_second = asv_rate * examples_per_second,
      robust_accuracy_per_second = if_else(
        split %in% c("test_attacked_seen", "test_attacked_unseen", "all_attacks"),
        accuracy / seconds_per_example,
        NA_real_
      ),
      asv_reduction_proxy_per_second = if_else(
        split %in% c("test_attacked_seen", "test_attacked_unseen", "all_attacks"),
        (1 - asv_rate) / seconds_per_example,
        NA_real_
      )
    ) |>
    sort_scenarios()

  write_inference_csv(quality_by_split, "14_quality_metrics_by_split.csv")
  write_inference_csv(inference_cost_quality, "15_inference_cost_quality_joined.csv")

  inference_cost_quality |>
    filter(split %in% c("test_clean", "all_attacks")) |>
    select(
      scenario_short,
      seed,
      split,
      seconds_per_example,
      examples_per_second,
      accuracy,
      asv_rate,
      invalid_output_rate,
      correct_examples_per_second,
      robust_accuracy_per_second,
      asv_reduction_proxy_per_second
    ) |>
    mutate(
      seconds_per_example = fmt_number(seconds_per_example, accuracy = 0.0001),
      examples_per_second = fmt_number(examples_per_second, accuracy = 0.001),
      accuracy = fmt_pct(accuracy),
      asv_rate = fmt_pct(asv_rate),
      invalid_output_rate = fmt_pct(invalid_output_rate),
      correct_examples_per_second = fmt_number(correct_examples_per_second, accuracy = 0.001),
      robust_accuracy_per_second = fmt_number(robust_accuracy_per_second, accuracy = 0.001),
      asv_reduction_proxy_per_second = fmt_number(asv_reduction_proxy_per_second, accuracy = 0.001)
    ) |>
    knitr::kable(
      caption = "Table 11 - Inference cost joined with task and attack metrics."
    )
} else {
  inference_cost_quality <- NULL
  cat("tables/descriptive/01_inference_compact.csv not found. Quality-cost join skipped.")
}
```

# Figures

## Split elapsed time by scenario

```{r fig-split-elapsed, fig.width=11, fig.height=6.5}
p_split_elapsed <- split_metadata |>
  ggplot(aes(x = scenario_short, y = elapsed_minutes, color = scenario_short)) +
  geom_point(aes(shape = factor(seed)), size = 3.1, position = position_jitter(width = 0.08, height = 0)) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.45,
    color = "black",
    linewidth = 0.35
  ) +
  facet_wrap(~ split_label, scales = "free_y") +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  labs(
    title = "Inference elapsed time by split",
    subtitle = "Points are individual seeds; black bars are scenario means.",
    x = "Scenario",
    y = "Elapsed time (minutes)",
    color = "Scenario",
    shape = "Seed"
  ) +
  base_theme()

print(p_split_elapsed)

save_inference_figure(
  p_split_elapsed,
  "01_inference_elapsed_time_by_split.png",
  width = 11,
  height = 6.5
)
```

## Seconds per example by split

```{r fig-seconds-per-example, fig.width=11, fig.height=6.5}
p_seconds_per_example <- split_metadata |>
  ggplot(aes(x = scenario_short, y = seconds_per_example, color = scenario_short)) +
  geom_point(aes(shape = factor(seed)), size = 3.1, position = position_jitter(width = 0.08, height = 0)) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.45,
    color = "black",
    linewidth = 0.35
  ) +
  facet_wrap(~ split_label, scales = "free_y") +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  labs(
    title = "Inference seconds per example",
    subtitle = "Lower values mean cheaper inference per example.",
    x = "Scenario",
    y = "Seconds per example",
    color = "Scenario",
    shape = "Seed"
  ) +
  base_theme()

print(p_seconds_per_example)

save_inference_figure(
  p_seconds_per_example,
  "02_inference_seconds_per_example_by_split.png",
  width = 11,
  height = 6.5
)
```

## Examples per second by split

```{r fig-examples-per-second, fig.width=11, fig.height=6.5}
p_examples_per_second <- split_metadata |>
  ggplot(aes(x = scenario_short, y = examples_per_second, color = scenario_short)) +
  geom_point(aes(shape = factor(seed)), size = 3.1, position = position_jitter(width = 0.08, height = 0)) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.45,
    color = "black",
    linewidth = 0.35
  ) +
  facet_wrap(~ split_label, scales = "free_y") +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  labs(
    title = "Inference throughput by split",
    subtitle = "Higher values mean more examples processed per second.",
    x = "Scenario",
    y = "Examples per second",
    color = "Scenario",
    shape = "Seed"
  ) +
  base_theme()

print(p_examples_per_second)

save_inference_figure(
  p_examples_per_second,
  "03_inference_examples_per_second_by_split.png",
  width = 11,
  height = 6.5
)
```

## Relative overhead vs C0

```{r fig-relative-overhead, fig.width=11, fig.height=6.5}
p_relative_overhead <- overhead_vs_c0 |>
  filter(split %in% c("test_clean", "test_attacked_seen", "test_attacked_unseen", "all_splits")) |>
  ggplot(aes(x = scenario_short, y = relative_seconds_per_example_vs_c0, fill = scenario_short)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey45") +
  geom_col(width = 0.72, show.legend = FALSE) +
  facet_wrap(~ split_label, scales = "free_y") +
  scale_fill_manual(values = scenario_palette, drop = FALSE) +
  labs(
    title = "Relative inference cost vs C0",
    subtitle = "Values above 1.0x mean higher seconds per example than C0.",
    x = "Scenario",
    y = "Relative seconds per example vs C0"
  ) +
  base_theme()

print(p_relative_overhead)

save_inference_figure(
  p_relative_overhead,
  "04_relative_inference_cost_vs_c0.png",
  width = 11,
  height = 6.5
)
```

## Run-level total inference time

```{r fig-run-total, fig.width=10.5, fig.height=6}
p_run_total <- run_metadata |>
  ggplot(aes(x = scenario_short, y = elapsed_minutes, color = scenario_short)) +
  geom_point(aes(shape = factor(seed)), size = 3.4, position = position_jitter(width = 0.08, height = 0)) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.45,
    color = "black",
    linewidth = 0.35
  ) +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  labs(
    title = "Run-level total inference time",
    subtitle = "Includes the complete scenario-seed inference run.",
    x = "Scenario",
    y = "Elapsed time (minutes)",
    color = "Scenario",
    shape = "Seed"
  ) +
  base_theme()

print(p_run_total)

save_inference_figure(
  p_run_total,
  "05_run_level_total_inference_time.png",
  width = 10.5,
  height = 6
)
```

## Run overhead

```{r fig-run-overhead, fig.width=10.5, fig.height=6}
p_run_overhead <- run_metadata |>
  ggplot(aes(x = scenario_short, y = run_overhead_rate, color = scenario_short)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey45") +
  geom_point(aes(shape = factor(seed)), size = 3.4, position = position_jitter(width = 0.08, height = 0)) +
  stat_summary(
    fun = mean,
    geom = "crossbar",
    width = 0.45,
    color = "black",
    linewidth = 0.35
  ) +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Run-level overhead",
    subtitle = "Overhead = run elapsed time minus sum of split elapsed times.",
    x = "Scenario",
    y = "Overhead rate",
    color = "Scenario",
    shape = "Seed"
  ) +
  base_theme()

print(p_run_overhead)

save_inference_figure(
  p_run_overhead,
  "06_run_level_inference_overhead.png",
  width = 10.5,
  height = 6
)
```

## Normalized null rate

```{r fig-null-rate, fig.width=11, fig.height=6.5}
p_null_rate <- split_summary_by_scenario |>
  ggplot(aes(x = scenario_short, y = normalized_null_rate, fill = scenario_short)) +
  geom_col(width = 0.72, show.legend = FALSE) +
  facet_wrap(~ split_label, scales = "free_y") +
  scale_fill_manual(values = scenario_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1)) +
  labs(
    title = "Normalized null rate by inference split",
    subtitle = "Higher values indicate more null normalized outputs during inference.",
    x = "Scenario",
    y = "Normalized null rate"
  ) +
  base_theme()

print(p_null_rate)

save_inference_figure(
  p_null_rate,
  "07_normalized_null_rate_by_split.png",
  width = 11,
  height = 6.5
)
```

## Inference timeline

```{r fig-inference-timeline, fig.width=11, fig.height=6}
timeline_data <- run_metadata |>
  mutate(
    run_label = paste0(scenario_short, " seed ", seed),
    run_label = fct_reorder(run_label, as.numeric(started_at))
  )

p_timeline <- timeline_data |>
  ggplot(aes(y = run_label, color = scenario_short)) +
  geom_segment(
    aes(x = started_at, xend = finished_at, yend = run_label),
    linewidth = 4,
    lineend = "round"
  ) +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  labs(
    title = "Inference run timeline",
    subtitle = "Wall-clock interval from started_at_utc to finished_at_utc.",
    x = "Time (UTC)",
    y = "Run",
    color = "Scenario"
  ) +
  base_theme()

print(p_timeline)

save_inference_figure(
  p_timeline,
  "08_inference_run_timeline.png",
  width = 11,
  height = 6
)
```

## Split timeline

```{r fig-split-timeline, fig.width=11, fig.height=7}
split_timeline_data <- split_metadata |>
  mutate(
    run_label = paste0(scenario_short, " seed ", seed),
    run_label = fct_reorder(run_label, as.numeric(started_at))
  )

p_split_timeline <- split_timeline_data |>
  ggplot(aes(y = run_label, color = split_label)) +
  geom_segment(
    aes(x = started_at, xend = finished_at, yend = run_label),
    linewidth = 4,
    lineend = "round"
  ) +
  labs(
    title = "Inference split timeline",
    subtitle = "Shows clean, attacked-seen, and attacked-unseen split durations.",
    x = "Time (UTC)",
    y = "Run",
    color = "Split"
  ) +
  base_theme()

print(p_split_timeline)

save_inference_figure(
  p_split_timeline,
  "09_inference_split_timeline.png",
  width = 11,
  height = 7
)
```

## Cost-quality figures

```{r fig-cost-quality, fig.width=10.5, fig.height=6.5}
if (!is.null(inference_cost_quality)) {
  cost_quality_all_attacks <- inference_cost_quality |>
    filter(split == "all_attacks")

  p_cost_quality <- cost_quality_all_attacks |>
    ggplot(aes(x = seconds_per_example, y = accuracy, color = scenario_short)) +
    geom_point(aes(shape = factor(seed)), size = 3.4) +
    scale_color_manual(values = scenario_palette, drop = FALSE) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(
      title = "Inference cost vs robust accuracy",
      subtitle = "Ideal region is upper-left: high robust accuracy and low seconds per example.",
      x = "Seconds per example",
      y = "Robust accuracy",
      color = "Scenario",
      shape = "Seed"
    ) +
    base_theme()

  print(p_cost_quality)

  save_inference_figure(
    p_cost_quality,
    "10_inference_cost_vs_robust_accuracy.png",
    width = 10.5,
    height = 6.5
  )

  p_cost_asv <- cost_quality_all_attacks |>
    ggplot(aes(x = seconds_per_example, y = asv_rate, color = scenario_short)) +
    geom_point(aes(shape = factor(seed)), size = 3.4) +
    scale_color_manual(values = scenario_palette, drop = FALSE) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(
      title = "Inference cost vs ASV",
      subtitle = "Ideal region is lower-left: low attack success and low seconds per example.",
      x = "Seconds per example",
      y = "ASV",
      color = "Scenario",
      shape = "Seed"
    ) +
    base_theme()

  print(p_cost_asv)

  save_inference_figure(
    p_cost_asv,
    "11_inference_cost_vs_asv.png",
    width = 10.5,
    height = 6.5
  )
}
```

# Paper-ready inference cost table

```{r paper-ready-table}
paper_inference_cost_table <- scenario_inference_summary |>
  filter(split %in% c("test_clean", "all_attacks", "all_splits")) |>
  select(
    scenario_short,
    split_label,
    prompt_strategy,
    n_runs,
    seeds,
    mean_rows,
    mean_elapsed_minutes,
    mean_seconds_per_example,
    mean_examples_per_second,
    normalized_null_rate,
    total_elapsed_hours,
    total_duration_label
  ) |>
  mutate(
    mean_rows = fmt_number(mean_rows, accuracy = 1),
    mean_elapsed_minutes = fmt_number(mean_elapsed_minutes),
    mean_seconds_per_example = fmt_number(mean_seconds_per_example, accuracy = 0.0001),
    mean_examples_per_second = fmt_number(mean_examples_per_second, accuracy = 0.001),
    normalized_null_rate = fmt_pct(normalized_null_rate),
    total_elapsed_hours = fmt_number(total_elapsed_hours)
  )

write_inference_csv(paper_inference_cost_table, "16_paper_ready_inference_cost_table.csv")

paper_inference_cost_table |>
  knitr::kable(
    caption = "Table 12 - Paper-ready inference cost table."
  )
```

# Methodological notes

```{r methodological-notes, results='asis'}
cat("
## How to interpret this report

Elapsed time is the wall-clock time reported by the inference metadata. Seconds per example is the main normalized inference-cost metric because the splits have different numbers of rows.

Examples per second is the reciprocal throughput view. Higher examples per second means faster inference.

Relative overhead vs C0 compares mean seconds per example for each scenario and split against the C0 base-model reference. Values above 1.0x indicate more expensive inference than C0.

C1 is useful for estimating the cost of the StruQ formatting strategy without adapter training. C2, C3, and C4 include trained adapter-based scenarios.

The quality-cost section is optional. It appears only when tables/descriptive/01_inference_compact.csv is available.
")
```

# Generated files

Tables are saved in:

```text
tables/inference_time_analysis/
```

Figures are saved in:

```text
figures/inference_time_analysis/
```

Main files:

```text
01_inference_experiment_metadata.csv
02_inference_result_files_from_summary.csv
03_inference_split_metadata_enriched.csv
04_inference_run_metadata_enriched.csv
05_split_summary_by_scenario.csv
06_aggregate_split_metadata.csv
07_scenario_inference_summary.csv
08_inference_overhead_vs_c0.csv
09_prompt_strategy_inference_summary.csv
10_inference_events_enriched.csv
11_inference_event_counts.csv
12_split_completion_events.csv
13_inference_error_files.csv
14_quality_metrics_by_split.csv
15_inference_cost_quality_joined.csv
16_paper_ready_inference_cost_table.csv

01_inference_elapsed_time_by_split.png
02_inference_seconds_per_example_by_split.png
03_inference_examples_per_second_by_split.png
04_relative_inference_cost_vs_c0.png
05_run_level_total_inference_time.png
06_run_level_inference_overhead.png
07_normalized_null_rate_by_split.png
08_inference_run_timeline.png
09_inference_split_timeline.png
10_inference_cost_vs_robust_accuracy.png
11_inference_cost_vs_asv.png
```
