---
title: "RPC-PI Experiment - Total Pipeline Runtime Analysis"
author: ""
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    theme: sandstone
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)

options(knitr.kable.NA = "-")

required_packages <- c(
  "tidyverse",
  "readr",
  "ggplot2",
  "scales",
  "stringr",
  "forcats",
  "fs",
  "here",
  "knitr",
  "jsonlite",
  "lubridate"
)

missing_packages <- required_packages[!required_packages %in% rownames(installed.packages())]
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

library(tidyverse)
library(readr)
library(ggplot2)
library(scales)
library(stringr)
library(forcats)
library(fs)
library(here)
library(knitr)
library(jsonlite)
library(lubridate)
```

# Objective

This R Markdown analyzes the **total runtime cost of the experimental pipeline**.

It combines logs from:

```text
training
inference
metrics
pairwise_metrics
statistical_analysis
error_analysis
```

The main questions are:

```text
How long did the complete pipeline take?
Which stage dominates the runtime?
How much time was spent in training, inference, pairwise/judge metrics, and analysis?
How many runs, rows, events, and errors are associated with each stage?
Where are the main pipeline bottlenecks?
```

This report is complementary to:

```text
14_training_time_analysis.Rmd
15_inference_time_analysis.Rmd
```

The focus here is not only model cost, but **end-to-end experiment cost**.

# Expected folder structure

This Rmd expects the logs extracted into one of these locations:

```text
data/logs/
logs/
```

If `logs.zip` exists in the project root or inside `data/`, this Rmd tries to extract it automatically to `data/`.

# Paths

```{r paths}
OUT_TABLE_DIR <- here::here("tables", "pipeline_runtime_analysis")
OUT_FIG_DIR <- here::here("figures", "pipeline_runtime_analysis")

fs::dir_create(OUT_TABLE_DIR)
fs::dir_create(OUT_FIG_DIR)

# Try to extract logs.zip automatically if needed.
if (!dir.exists(here::here("data", "logs")) && file.exists(here::here("logs.zip"))) {
  dir.create(here::here("data"), showWarnings = FALSE)
  unzip(here::here("logs.zip"), exdir = here::here("data"))
}

if (!dir.exists(here::here("data", "logs")) && file.exists(here::here("data", "logs.zip"))) {
  dir.create(here::here("data"), showWarnings = FALSE)
  unzip(here::here("data", "logs.zip"), exdir = here::here("data"))
}

candidate_logs_dirs <- c(
  here::here("data", "logs"),
  here::here("logs")
)

LOGS_DIR <- candidate_logs_dirs[dir.exists(candidate_logs_dirs)][1]

if (is.na(LOGS_DIR)) {
  stop(
    "Could not find logs directory. Expected one of:\n",
    paste(candidate_logs_dirs, collapse = "\n"),
    "\nOr place logs.zip in the project root or data/."
  )
}
```

# Helper functions

```{r helpers}
`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0) y else x
}

scenario_order <- c(
  "c0_base",
  "c1_struq_format_only",
  "c2_struq_sft",
  "c3_secalign_dpo",
  "c4_ih_sft"
)

scenario_labels <- tibble(
  scenario_id = scenario_order,
  scenario_short = c("C0", "C1", "C2", "C3", "C4"),
  scenario_label = c(
    "C0 - Base model",
    "C1 - StruQ format-only",
    "C2 - StruQ-like SFT",
    "C3 - SecAlign-like DPO",
    "C4 - Instruction-Hierarchy-like SFT"
  )
)

scenario_palette <- c(
  C0 = "#7F7F7F",
  C1 = "#D55E00",
  C2 = "#0072B2",
  C3 = "#CC79A7",
  C4 = "#009E73"
)

stage_order <- c(
  "training",
  "inference",
  "metrics",
  "pairwise_metrics",
  "statistical_analysis",
  "error_analysis"
)

stage_labels <- c(
  training = "Training",
  inference = "Inference",
  metrics = "Metrics",
  pairwise_metrics = "Pairwise metrics",
  statistical_analysis = "Statistical analysis",
  error_analysis = "Error analysis"
)

stage_palette <- c(
  training = "#0072B2",
  inference = "#009E73",
  metrics = "#E69F00",
  pairwise_metrics = "#CC79A7",
  statistical_analysis = "#56B4E9",
  error_analysis = "#D55E00"
)

sort_stages <- function(df) {
  df |>
    mutate(
      stage = factor(as.character(stage), levels = stage_order),
      stage_label = factor(
        ifelse(is.na(stage_labels[as.character(stage)]), as.character(stage), stage_labels[as.character(stage)]),
        levels = stage_labels[stage_order]
      )
    ) |>
    arrange(stage)
}

sort_scenarios <- function(df) {
  df |>
    mutate(
      scenario_id = factor(as.character(scenario_id), levels = scenario_order),
      scenario_short = factor(as.character(scenario_short), levels = c("C0", "C1", "C2", "C3", "C4"))
    ) |>
    arrange(scenario_id)
}

add_scenario_info <- function(df) {
  if (!("scenario_id" %in% names(df))) {
    df$scenario_id <- NA_character_
  }

  out <- df |>
    mutate(scenario_id = as.character(scenario_id)) |>
    left_join(scenario_labels, by = "scenario_id", suffix = c("", "_lookup"))

  if (!("scenario_label" %in% names(out))) {
    out$scenario_label <- out$scenario_id
  }

  if ("scenario_label_lookup" %in% names(out)) {
    out <- out |>
      mutate(
        scenario_label = coalesce(
          as.character(scenario_label),
          as.character(scenario_label_lookup)
        )
      )
  }

  if (!("scenario_short" %in% names(out))) {
    out$scenario_short <- NA_character_
  }

  if ("scenario_short_lookup" %in% names(out)) {
    out <- out |>
      mutate(
        scenario_short = coalesce(
          as.character(scenario_short),
          as.character(scenario_short_lookup)
        )
      )
  }

  out |>
    mutate(
      scenario_short = ifelse(
        is.na(scenario_short),
        recode(
          as.character(scenario_id),
          c0_base = "C0",
          c1_struq_format_only = "C1",
          c2_struq_sft = "C2",
          c3_secalign_dpo = "C3",
          c4_ih_sft = "C4",
          .default = as.character(scenario_id)
        ),
        scenario_short
      )
    ) |>
    select(-any_of(c("scenario_label_lookup", "scenario_short_lookup"))) |>
    sort_scenarios()
}

parse_time_utc <- function(x) {
  lubridate::ymd_hms(x, tz = "UTC", quiet = TRUE)
}

safe_mean <- function(x) {
  if (sum(!is.na(x)) == 0) return(NA_real_)
  mean(x, na.rm = TRUE)
}

safe_sd <- function(x) {
  if (sum(!is.na(x)) <= 1) return(NA_real_)
  sd(x, na.rm = TRUE)
}

fmt_duration <- function(seconds) {
  ifelse(
    is.na(seconds),
    NA_character_,
    paste0(
      floor(seconds / 3600), "h ",
      floor((seconds %% 3600) / 60), "m ",
      round(seconds %% 60), "s"
    )
  )
}

fmt_number <- function(x, accuracy = 0.01) {
  scales::number(x, accuracy = accuracy, big.mark = ",")
}

fmt_pct <- function(x, accuracy = 0.1) {
  scales::percent(x, accuracy = accuracy)
}

base_theme <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      legend.position = "bottom",
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(color = "grey30"),
      panel.grid.minor = element_blank(),
      axis.title = element_text(face = "bold")
    )
}

sanitize_for_csv <- function(df) {
  df <- as_tibble(df)
  n_rows <- nrow(df)

  value_to_text <- function(value) {
    if (is.null(value) || length(value) == 0) {
      return(NA_character_)
    }

    tryCatch(
      as.character(jsonlite::toJSON(value, auto_unbox = TRUE, null = "null")),
      error = function(e) paste(capture.output(str(value)), collapse = " ")
    )
  }

  for (nm in names(df)) {
    col <- df[[nm]]

    if (is.data.frame(col)) {
      if (nrow(col) == n_rows) {
        df[[nm]] <- purrr::map_chr(
          seq_len(n_rows),
          function(i) value_to_text(as.list(col[i, , drop = FALSE]))
        )
      } else {
        df[[nm]] <- rep(value_to_text(col), n_rows)
      }
    } else if (is.matrix(col) || is.array(col)) {
      mat <- as.matrix(col)

      if (nrow(mat) == n_rows) {
        df[[nm]] <- apply(
          mat,
          1,
          function(row) paste(as.character(row), collapse = " | ")
        )
      } else {
        df[[nm]] <- rep(value_to_text(mat), n_rows)
      }
    } else if (is.list(col)) {
      if (length(col) == n_rows) {
        df[[nm]] <- purrr::map_chr(col, value_to_text)
      } else if (length(col) == 1) {
        df[[nm]] <- rep(value_to_text(col[[1]]), n_rows)
      } else {
        df[[nm]] <- rep(value_to_text(col), n_rows)
      }
    }
  }

  df
}

write_pipeline_csv <- function(df, filename) {
  safe_df <- sanitize_for_csv(df)
  readr::write_csv(safe_df, file.path(OUT_TABLE_DIR, filename))
  invisible(df)
}

save_pipeline_figure <- function(plot, filename, width = 10, height = 6) {
  ggsave(
    file.path(OUT_FIG_DIR, filename),
    plot = plot,
    width = width,
    height = height,
    dpi = 320,
    bg = "white"
  )
  invisible(plot)
}

replace_null_columns <- function(x) {
  if (!is.list(x)) {
    return(x)
  }

  n <- 1
  non_null_lengths <- purrr::map_int(
    x,
    function(value) {
      if (is.null(value)) {
        0L
      } else if (is.data.frame(value)) {
        nrow(value)
      } else {
        length(value)
      }
    }
  )

  if (any(non_null_lengths > 0)) {
    n <- max(non_null_lengths)
  }

  purrr::map(
    x,
    function(value) {
      if (is.null(value)) {
        rep(NA_character_, n)
      } else {
        value
      }
    }
  )
}

as_tibble_json_safe <- function(x) {
  if (is.null(x)) {
    return(tibble())
  }

  if (is.data.frame(x)) {
    x <- replace_null_columns(x)
    return(tibble::as_tibble(x))
  }

  if (is.list(x)) {
    x <- replace_null_columns(x)
    return(tibble::as_tibble(x))
  }

  tibble(value = x)
}

read_jsonl_robust <- function(path, stage_name, bad_lines_filename = "00_bad_jsonl_lines.csv") {
  if (!file.exists(path)) {
    return(tibble())
  }

  lines <- readLines(path, warn = FALSE, encoding = "UTF-8")

  if (length(lines) == 0) {
    return(tibble())
  }

  parsed <- vector("list", length(lines))
  bad_lines <- list()

  for (i in seq_along(lines)) {
    line <- lines[[i]]

    if (is.na(line) || trimws(line) == "") {
      next
    }

    obj <- tryCatch(
      jsonlite::fromJSON(line, flatten = TRUE),
      error = function(e) {
        bad_lines[[length(bad_lines) + 1]] <<- tibble(
          stage = stage_name,
          line_number = i,
          error = conditionMessage(e),
          line_prefix = substr(line, 1, 250)
        )
        NULL
      }
    )

    if (!is.null(obj)) {
      parsed[[i]] <- as_tibble_json_safe(obj)
    }
  }

  parsed <- parsed[!vapply(parsed, is.null, logical(1))]
  bad_lines_df <- dplyr::bind_rows(bad_lines)

  if (nrow(bad_lines_df) > 0) {
    readr::write_csv(
      bad_lines_df,
      file.path(OUT_TABLE_DIR, bad_lines_filename)
    )
  }

  if (length(parsed) == 0) {
    return(tibble())
  }

  dplyr::bind_rows(parsed) |>
    mutate(stage = stage_name, .before = 1)
}

get_scalar <- function(obj, field, default = NA) {
  value <- obj[[field]]

  if (is.null(value) || length(value) == 0) {
    return(default)
  }

  if (is.data.frame(value)) {
    if (nrow(value) == 0 || ncol(value) == 0) return(default)
    return(value[[1]][[1]])
  }

  if (is.list(value) && length(value) == 1 && !is.null(value[[1]])) {
    return(value[[1]])
  }

  value[[1]]
}

get_nested_scalar <- function(obj, parent, child, default = NA) {
  flat_name <- paste0(parent, ".", child)

  flat_value <- obj[[flat_name]]

  if (!is.null(flat_value) && length(flat_value) > 0) {
    return(flat_value[[1]])
  }

  parent_value <- obj[[parent]]

  if (!is.null(parent_value)) {
    if (is.data.frame(parent_value) && child %in% names(parent_value)) {
      return(parent_value[[child]][[1]])
    }

    if (is.list(parent_value) && !is.null(parent_value[[child]])) {
      return(parent_value[[child]][[1]])
    }
  }

  default
}

read_json_safe <- function(path) {
  tryCatch(
    jsonlite::fromJSON(path, flatten = TRUE),
    error = function(e) NULL
  )
}
```

# Discover log files

```{r discover-logs}
stage_paths <- tibble(
  stage = stage_order,
  stage_label = stage_labels[stage_order],
  stage_dir = file.path(LOGS_DIR, stage_order)
) |>
  mutate(
    dir_exists = dir.exists(stage_dir),
    event_path = case_when(
      stage == "training" ~ file.path(stage_dir, "04_run_training_events.jsonl"),
      stage == "inference" ~ file.path(stage_dir, "05_run_inference_events.jsonl"),
      stage == "metrics" ~ file.path(stage_dir, "06_compute_metrics_events.jsonl"),
      stage == "pairwise_metrics" ~ file.path(stage_dir, "07_pairwise_and_injected_metrics_events.jsonl"),
      stage == "statistical_analysis" ~ file.path(stage_dir, "08_statistical_analysis_events.jsonl"),
      stage == "error_analysis" ~ file.path(stage_dir, "09_error_analysis_events.jsonl"),
      TRUE ~ NA_character_
    ),
    summary_path = case_when(
      stage == "training" ~ file.path(stage_dir, "04_run_training_summary.json"),
      stage == "inference" ~ file.path(stage_dir, "05_run_inference_summary.json"),
      stage == "metrics" ~ file.path(stage_dir, "06_compute_metrics_summary_full.json"),
      stage == "pairwise_metrics" ~ file.path(stage_dir, "07_pairwise_and_injected_metrics_summary.json"),
      stage == "statistical_analysis" ~ file.path(stage_dir, "08_statistical_analysis_summary_full.json"),
      stage == "error_analysis" ~ file.path(stage_dir, "09_error_analysis_summary.json"),
      TRUE ~ NA_character_
    ),
    event_exists = file.exists(event_path),
    summary_exists = file.exists(summary_path)
  ) |>
  sort_stages()

write_pipeline_csv(stage_paths, "01_stage_paths.csv")

stage_paths |>
  knitr::kable(
    caption = "Table 1 - Discovered pipeline log files."
  )
```

# Load all stage events

```{r load-events}
all_events <- purrr::map2_dfr(
  stage_paths$event_path,
  stage_paths$stage,
  function(path, stage_name) {
    read_jsonl_robust(
      path = path,
      stage_name = stage_name,
      bad_lines_filename = "00_bad_pipeline_event_lines.csv"
    )
  }
)

if (nrow(all_events) == 0) {
  stop("No pipeline events could be loaded.")
}

events_raw_names <- names(all_events)

if (!("scenario_id" %in% events_raw_names) && "scenario" %in% events_raw_names) {
  all_events <- all_events |>
    rename(scenario_id = scenario)
} else if ("scenario" %in% events_raw_names) {
  all_events <- all_events |>
    mutate(
      scenario_id = coalesce(
        as.character(scenario_id),
        as.character(scenario)
      )
    )
}

if (!("seed" %in% names(all_events))) {
  all_events$seed <- NA_integer_
}

if (!("event_type" %in% names(all_events))) {
  all_events$event_type <- NA_character_
}

if (!("timestamp_utc" %in% names(all_events))) {
  stop("Events do not contain timestamp_utc.")
}

if (!("run_mode" %in% names(all_events))) {
  all_events$run_mode <- NA_character_
}

if (!("rows" %in% names(all_events))) {
  all_events$rows <- NA_real_
}

if (!("elapsed_seconds" %in% names(all_events))) {
  all_events$elapsed_seconds <- NA_real_
}

all_events <- all_events |>
  mutate(
    stage = as.character(stage),
    event_type = as.character(event_type),
    timestamp = parse_time_utc(timestamp_utc),
    scenario_id = as.character(scenario_id),
    seed = suppressWarnings(as.integer(seed)),
    run_mode = as.character(run_mode),
    rows = suppressWarnings(as.numeric(rows)),
    elapsed_seconds = suppressWarnings(as.numeric(elapsed_seconds))
  ) |>
  add_scenario_info() |>
  sort_stages()

write_pipeline_csv(all_events, "02_all_pipeline_events.csv")

all_events |>
  count(stage_label, event_type, name = "n_events") |>
  arrange(stage_label, desc(n_events)) |>
  knitr::kable(
    caption = "Table 2 - Event counts by stage."
  )
```

# Stage timelines from event logs

```{r stage-timelines}
stage_event_summary <- all_events |>
  group_by(stage, stage_label) |>
  summarise(
    n_events = n(),
    first_event_utc = min(timestamp, na.rm = TRUE),
    last_event_utc = max(timestamp, na.rm = TRUE),
    observed_stage_seconds = as.numeric(difftime(last_event_utc, first_event_utc, units = "secs")),
    observed_stage_minutes = observed_stage_seconds / 60,
    observed_stage_hours = observed_stage_seconds / 3600,
    event_elapsed_seconds_sum = sum(elapsed_seconds, na.rm = TRUE),
    event_rows_sum = sum(rows, na.rm = TRUE),
    n_event_rows_with_elapsed = sum(!is.na(elapsed_seconds)),
    n_event_rows_with_rows = sum(!is.na(rows)),
    .groups = "drop"
  ) |>
  sort_stages()

pipeline_start <- min(stage_event_summary$first_event_utc, na.rm = TRUE)
pipeline_end <- max(stage_event_summary$last_event_utc, na.rm = TRUE)
pipeline_wall_clock_seconds <- as.numeric(difftime(pipeline_end, pipeline_start, units = "secs"))

stage_event_summary <- stage_event_summary |>
  mutate(
    share_of_wall_clock = observed_stage_seconds / pipeline_wall_clock_seconds,
    observed_duration_label = fmt_duration(observed_stage_seconds)
  )

write_pipeline_csv(stage_event_summary, "03_stage_event_timeline_summary.csv")

stage_event_summary |>
  select(
    stage_label,
    n_events,
    first_event_utc,
    last_event_utc,
    observed_stage_minutes,
    observed_stage_hours,
    share_of_wall_clock,
    event_elapsed_seconds_sum,
    event_rows_sum,
    observed_duration_label
  ) |>
  mutate(
    observed_stage_minutes = fmt_number(observed_stage_minutes),
    observed_stage_hours = fmt_number(observed_stage_hours),
    share_of_wall_clock = fmt_pct(share_of_wall_clock),
    event_elapsed_seconds_sum = fmt_number(event_elapsed_seconds_sum),
    event_rows_sum = fmt_number(event_rows_sum, accuracy = 1)
  ) |>
  knitr::kable(
    caption = "Table 3 - Stage timeline summary based on event timestamps."
  )
```

# Load summary JSON files

```{r load-summaries}
summary_records <- purrr::pmap_dfr(
  stage_paths |>
    select(stage, stage_label, summary_path, summary_exists),
  function(stage, stage_label, summary_path, summary_exists) {
    if (!isTRUE(summary_exists)) {
      return(tibble(
        stage = stage,
        stage_label = stage_label,
        summary_exists = FALSE
      ))
    }

    obj <- read_json_safe(summary_path)

    if (is.null(obj)) {
      return(tibble(
        stage = stage,
        stage_label = stage_label,
        summary_exists = TRUE,
        summary_parse_ok = FALSE
      ))
    }

    counts <- obj$counts %||% list()
    outputs <- obj$outputs %||% list()

    tibble(
      stage = stage,
      stage_label = stage_label,
      summary_exists = TRUE,
      summary_parse_ok = TRUE,
      notebook = as.character(obj$notebook %||% NA_character_),
      created_at_utc = as.character(obj$created_at_utc %||% NA_character_),
      run_mode = as.character(obj$run_mode %||% NA_character_),
      base_model_id = as.character(obj$base_model_id %||% obj$base_model %||% NA_character_),
      judge_model_id = as.character(obj$judge_model_id %||% NA_character_),
      num_inference_rows = as.numeric(obj$num_inference_rows %||% NA_real_),
      total_inference_rows = as.numeric(obj$total_inference_rows %||% NA_real_),
      num_scenarios = as.numeric(obj$num_scenarios %||% NA_real_),
      num_scenario_seed_runs = as.numeric(obj$num_scenario_seed_runs %||% NA_real_),
      training_results_count = length(obj$training_results %||% list()),
      inference_results_count = length(obj$inference_results %||% list()),
      result_files_count = length(obj$result_files %||% list()),
      input_files_count = length(obj$input_files %||% list()),
      generated_files_count = length(obj$generated_files %||% list()),
      metrics_computed_count = length(obj$metrics_computed %||% list()),
      metrics_available_count = length(obj$metrics_available %||% list()),
      outputs_count = length(outputs),
      counts_injected_only_rows = as.numeric(counts$injected_only_rows %||% NA_real_),
      counts_mr_paired_rows = as.numeric(counts$mr_paired_rows %||% NA_real_),
      counts_win_rate_pairs = as.numeric(counts$win_rate_pairs %||% NA_real_),
      counts_win_rate_judgments = as.numeric(counts$win_rate_judgments %||% NA_real_),
      counts_unified_run_rows = as.numeric(counts$unified_run_rows %||% NA_real_),
      counts_scenario_metric_summary_rows = as.numeric(counts$scenario_metric_summary_rows %||% NA_real_),
      coverage_rows = as.numeric(obj$coverage_rows %||% NA_real_),
      failure_taxonomy_rows = as.numeric(obj$failure_taxonomy_rows %||% NA_real_),
      paired_vs_c0_rows = as.numeric(obj$paired_vs_c0_rows %||% NA_real_),
      invalid_output_rows = as.numeric(obj$invalid_output_rows %||% NA_real_),
      summary_path = summary_path
    )
  }
) |>
  sort_stages()

write_pipeline_csv(summary_records, "04_stage_summary_json_records.csv")

summary_records |>
  select(
    stage_label,
    notebook,
    created_at_utc,
    run_mode,
    base_model_id,
    judge_model_id,
    training_results_count,
    inference_results_count,
    result_files_count,
    input_files_count,
    generated_files_count,
    metrics_computed_count,
    outputs_count
  ) |>
  knitr::kable(
    caption = "Table 4 - Summary JSON overview by stage."
  )
```

# Load run metadata files

```{r load-run-metadata}
all_run_metadata_files <- fs::dir_ls(LOGS_DIR, recurse = TRUE, regexp = "run_metadata\\.json$")

read_generic_run_metadata <- function(path) {
  obj <- read_json_safe(path)

  if (is.null(obj)) {
    return(tibble(
      metadata_path = as.character(path),
      parse_ok = FALSE
    ))
  }

  path_chr <- as.character(path)

  stage <- case_when(
    str_detect(path_chr, "[/\\\\]training[/\\\\]") ~ "training",
    str_detect(path_chr, "[/\\\\]inference[/\\\\]") ~ "inference",
    str_detect(path_chr, "[/\\\\]pairwise_metrics[/\\\\]") ~ "pairwise_metrics",
    TRUE ~ NA_character_
  )

  substage <- case_when(
    str_detect(path_chr, "[/\\\\]pairwise_metrics[/\\\\].*[/\\\\]injected_only[/\\\\]") ~ "injected_only",
    str_detect(path_chr, "[/\\\\]pairwise_metrics[/\\\\].*[/\\\\]win_rate[/\\\\]") ~ "win_rate",
    TRUE ~ stage
  )

  scenario_id <- as.character(
    get_scalar(
      obj,
      "scenario_id",
      get_scalar(
        obj,
        "scenario",
        get_scalar(obj, "defended_scenario_id", NA_character_)
      )
    )
  )

  seed <- suppressWarnings(as.integer(
    get_scalar(
      obj,
      "seed",
      get_scalar(obj, "defended_seed", NA_integer_)
    )
  ))

  rows <- suppressWarnings(as.numeric(
    get_scalar(
      obj,
      "rows",
      get_scalar(
        obj,
        "train_rows",
        NA_real_
      )
    )
  ))

  train_rows <- suppressWarnings(as.numeric(get_scalar(obj, "train_rows", NA_real_)))
  validation_rows <- suppressWarnings(as.numeric(get_scalar(obj, "validation_rows", NA_real_)))

  splits <- obj$splits
  split_count <- if (is.null(splits)) {
    NA_integer_
  } else if (is.data.frame(splits)) {
    nrow(splits)
  } else if (is.list(splits)) {
    length(splits)
  } else {
    NA_integer_
  }

  split_rows_sum <- NA_real_
  split_elapsed_seconds_sum <- NA_real_

  if (!is.null(splits)) {
    split_tbl <- tryCatch(as_tibble_json_safe(splits), error = function(e) tibble())

    if (nrow(split_tbl) > 0) {
      if ("rows" %in% names(split_tbl)) {
        split_rows_sum <- sum(as.numeric(split_tbl$rows), na.rm = TRUE)
      }

      if ("elapsed_seconds" %in% names(split_tbl)) {
        split_elapsed_seconds_sum <- sum(as.numeric(split_tbl$elapsed_seconds), na.rm = TRUE)
      }
    }
  }

  tibble(
    stage = stage,
    substage = substage,
    scenario_id = scenario_id,
    seed = seed,
    status = as.character(get_scalar(obj, "status", NA_character_)),
    method = as.character(get_scalar(obj, "method", NA_character_)),
    run_mode = as.character(get_scalar(obj, "run_mode", NA_character_)),
    eval_split = as.character(get_scalar(obj, "eval_split", NA_character_)),
    started_at_utc = as.character(get_scalar(obj, "started_at_utc", NA_character_)),
    finished_at_utc = as.character(get_scalar(obj, "finished_at_utc", NA_character_)),
    started_at = parse_time_utc(started_at_utc),
    finished_at = parse_time_utc(finished_at_utc),
    elapsed_seconds = suppressWarnings(as.numeric(get_scalar(obj, "elapsed_seconds", NA_real_))),
    rows = rows,
    train_rows = train_rows,
    validation_rows = validation_rows,
    split_count = split_count,
    split_rows_sum = split_rows_sum,
    split_elapsed_seconds_sum = split_elapsed_seconds_sum,
    metadata_path = path_chr,
    parse_ok = TRUE
  )
}

run_metadata <- purrr::map_dfr(
  all_run_metadata_files,
  read_generic_run_metadata
) |>
  add_scenario_info() |>
  mutate(
    stage = as.character(stage),
    substage = as.character(substage),
    elapsed_minutes = elapsed_seconds / 60,
    elapsed_hours = elapsed_seconds / 3600,
    seconds_per_row = case_when(
      !is.na(split_rows_sum) & split_rows_sum > 0 ~ elapsed_seconds / split_rows_sum,
      !is.na(rows) & rows > 0 ~ elapsed_seconds / rows,
      TRUE ~ NA_real_
    ),
    rows_per_second = case_when(
      !is.na(split_rows_sum) & elapsed_seconds > 0 ~ split_rows_sum / elapsed_seconds,
      !is.na(rows) & elapsed_seconds > 0 ~ rows / elapsed_seconds,
      TRUE ~ NA_real_
    )
  ) |>
  sort_stages() |>
  sort_scenarios()

write_pipeline_csv(run_metadata, "05_all_run_metadata.csv")

run_metadata |>
  count(stage_label, substage, scenario_short, name = "n_run_metadata_files") |>
  arrange(stage_label, substage, scenario_short) |>
  knitr::kable(
    caption = "Table 5 - Run metadata files by stage, substage, and scenario."
  )
```

# Stage cost from run metadata

```{r stage-cost-from-runs}
stage_run_cost <- run_metadata |>
  filter(parse_ok, !is.na(stage)) |>
  group_by(stage, stage_label, substage) |>
  summarise(
    n_run_metadata_files = n(),
    n_completed_like = sum(str_detect(status, "completed"), na.rm = TRUE),
    n_error_like = sum(str_detect(status, "error|failed"), na.rm = TRUE),
    scenarios = paste(sort(unique(na.omit(as.character(scenario_short)))), collapse = ", "),
    seeds = paste(sort(unique(na.omit(seed))), collapse = ", "),
    total_run_elapsed_seconds = sum(elapsed_seconds, na.rm = TRUE),
    mean_run_elapsed_seconds = mean(elapsed_seconds, na.rm = TRUE),
    sd_run_elapsed_seconds = safe_sd(elapsed_seconds),
    total_rows_reported = sum(rows, na.rm = TRUE),
    total_split_rows = sum(split_rows_sum, na.rm = TRUE),
    total_train_rows = sum(train_rows, na.rm = TRUE),
    total_validation_rows = sum(validation_rows, na.rm = TRUE),
    total_split_elapsed_seconds = sum(split_elapsed_seconds_sum, na.rm = TRUE),
    mean_seconds_per_row = safe_mean(seconds_per_row),
    mean_rows_per_second = safe_mean(rows_per_second),
    .groups = "drop"
  ) |>
  mutate(
    total_run_elapsed_minutes = total_run_elapsed_seconds / 60,
    total_run_elapsed_hours = total_run_elapsed_seconds / 3600,
    mean_run_elapsed_minutes = mean_run_elapsed_seconds / 60,
    total_run_duration_label = fmt_duration(total_run_elapsed_seconds),
    mean_run_duration_label = fmt_duration(mean_run_elapsed_seconds)
  ) |>
  sort_stages()

write_pipeline_csv(stage_run_cost, "06_stage_cost_from_run_metadata.csv")

stage_run_cost |>
  select(
    stage_label,
    substage,
    n_run_metadata_files,
    n_completed_like,
    n_error_like,
    scenarios,
    total_run_elapsed_hours,
    mean_run_elapsed_minutes,
    total_rows_reported,
    total_split_rows,
    total_train_rows,
    total_validation_rows,
    mean_seconds_per_row,
    mean_rows_per_second,
    total_run_duration_label
  ) |>
  mutate(
    total_run_elapsed_hours = fmt_number(total_run_elapsed_hours),
    mean_run_elapsed_minutes = fmt_number(mean_run_elapsed_minutes),
    total_rows_reported = fmt_number(total_rows_reported, accuracy = 1),
    total_split_rows = fmt_number(total_split_rows, accuracy = 1),
    total_train_rows = fmt_number(total_train_rows, accuracy = 1),
    total_validation_rows = fmt_number(total_validation_rows, accuracy = 1),
    mean_seconds_per_row = fmt_number(mean_seconds_per_row, accuracy = 0.0001),
    mean_rows_per_second = fmt_number(mean_rows_per_second, accuracy = 0.001)
  ) |>
  knitr::kable(
    caption = "Table 6 - Stage cost from run metadata."
  )
```

# Unified stage cost summary

```{r unified-stage-cost}
stage_run_cost_total <- stage_run_cost |>
  group_by(stage, stage_label) |>
  summarise(
    run_metadata_elapsed_seconds = sum(total_run_elapsed_seconds, na.rm = TRUE),
    run_metadata_files = sum(n_run_metadata_files, na.rm = TRUE),
    run_metadata_completed_like = sum(n_completed_like, na.rm = TRUE),
    run_metadata_error_like = sum(n_error_like, na.rm = TRUE),
    run_metadata_rows = sum(total_rows_reported, na.rm = TRUE),
    run_metadata_split_rows = sum(total_split_rows, na.rm = TRUE),
    run_metadata_train_rows = sum(total_train_rows, na.rm = TRUE),
    .groups = "drop"
  )

unified_stage_cost <- stage_event_summary |>
  select(
    stage,
    stage_label,
    n_events,
    first_event_utc,
    last_event_utc,
    observed_stage_seconds,
    event_elapsed_seconds_sum,
    event_rows_sum
  ) |>
  full_join(
    stage_run_cost_total,
    by = c("stage", "stage_label")
  ) |>
  full_join(
    summary_records |>
      select(
        stage,
        stage_label,
        notebook,
        created_at_utc,
        run_mode,
        training_results_count,
        inference_results_count,
        result_files_count,
        input_files_count,
        generated_files_count,
        metrics_computed_count,
        outputs_count,
        num_inference_rows,
        total_inference_rows,
        counts_injected_only_rows,
        counts_mr_paired_rows,
        counts_win_rate_pairs,
        counts_win_rate_judgments,
        paired_vs_c0_rows,
        invalid_output_rows
      ),
    by = c("stage", "stage_label")
  ) |>
  mutate(
    primary_cost_seconds = case_when(
      !is.na(run_metadata_elapsed_seconds) & run_metadata_elapsed_seconds > 0 ~ run_metadata_elapsed_seconds,
      !is.na(event_elapsed_seconds_sum) & event_elapsed_seconds_sum > 0 ~ event_elapsed_seconds_sum,
      TRUE ~ observed_stage_seconds
    ),
    primary_cost_source = case_when(
      !is.na(run_metadata_elapsed_seconds) & run_metadata_elapsed_seconds > 0 ~ "run_metadata_elapsed_sum",
      !is.na(event_elapsed_seconds_sum) & event_elapsed_seconds_sum > 0 ~ "event_elapsed_sum",
      TRUE ~ "event_timestamp_span"
    ),
    primary_cost_minutes = primary_cost_seconds / 60,
    primary_cost_hours = primary_cost_seconds / 3600,
    observed_stage_minutes = observed_stage_seconds / 60,
    observed_stage_hours = observed_stage_seconds / 3600,
    run_metadata_elapsed_hours = run_metadata_elapsed_seconds / 3600,
    primary_cost_label = fmt_duration(primary_cost_seconds)
  ) |>
  mutate(
    share_of_primary_cost = primary_cost_seconds / sum(primary_cost_seconds, na.rm = TRUE),
    share_of_observed_stage_span = observed_stage_seconds / sum(observed_stage_seconds, na.rm = TRUE)
  ) |>
  sort_stages()

pipeline_summary <- tibble(
  pipeline_start_utc = pipeline_start,
  pipeline_end_utc = pipeline_end,
  pipeline_wall_clock_seconds = pipeline_wall_clock_seconds,
  pipeline_wall_clock_hours = pipeline_wall_clock_seconds / 3600,
  sum_observed_stage_seconds = sum(unified_stage_cost$observed_stage_seconds, na.rm = TRUE),
  sum_observed_stage_hours = sum_observed_stage_seconds / 3600,
  sum_primary_cost_seconds = sum(unified_stage_cost$primary_cost_seconds, na.rm = TRUE),
  sum_primary_cost_hours = sum_primary_cost_seconds / 3600,
  active_stage_share_of_wall_clock = sum_observed_stage_seconds / pipeline_wall_clock_seconds,
  primary_cost_share_of_wall_clock = sum_primary_cost_seconds / pipeline_wall_clock_seconds,
  pipeline_duration_label = fmt_duration(pipeline_wall_clock_seconds),
  sum_primary_cost_label = fmt_duration(sum_primary_cost_seconds)
)

write_pipeline_csv(unified_stage_cost, "07_unified_stage_cost_summary.csv")
write_pipeline_csv(pipeline_summary, "08_pipeline_total_summary.csv")

pipeline_summary |>
  mutate(
    pipeline_wall_clock_hours = fmt_number(pipeline_wall_clock_hours),
    sum_observed_stage_hours = fmt_number(sum_observed_stage_hours),
    sum_primary_cost_hours = fmt_number(sum_primary_cost_hours),
    active_stage_share_of_wall_clock = fmt_pct(active_stage_share_of_wall_clock),
    primary_cost_share_of_wall_clock = fmt_pct(primary_cost_share_of_wall_clock)
  ) |>
  knitr::kable(
    caption = "Table 7 - Total pipeline runtime summary."
  )

unified_stage_cost |>
  select(
    stage_label,
    notebook,
    n_events,
    run_metadata_files,
    first_event_utc,
    last_event_utc,
    observed_stage_hours,
    primary_cost_hours,
    primary_cost_source,
    share_of_primary_cost,
    event_rows_sum,
    run_metadata_rows,
    run_metadata_split_rows,
    primary_cost_label
  ) |>
  mutate(
    observed_stage_hours = fmt_number(observed_stage_hours),
    primary_cost_hours = fmt_number(primary_cost_hours),
    share_of_primary_cost = fmt_pct(share_of_primary_cost),
    event_rows_sum = fmt_number(event_rows_sum, accuracy = 1),
    run_metadata_rows = fmt_number(run_metadata_rows, accuracy = 1),
    run_metadata_split_rows = fmt_number(run_metadata_split_rows, accuracy = 1)
  ) |>
  knitr::kable(
    caption = "Table 8 - Unified stage cost summary."
  )
```

# Scenario-level pipeline cost

```{r scenario-level-pipeline-cost}
scenario_stage_cost <- run_metadata |>
  filter(parse_ok, !is.na(stage), !is.na(scenario_id)) |>
  group_by(stage, stage_label, substage, scenario_id, scenario_short, scenario_label) |>
  summarise(
    n_runs = n(),
    seeds = paste(sort(unique(na.omit(seed))), collapse = ", "),
    total_elapsed_seconds = sum(elapsed_seconds, na.rm = TRUE),
    mean_elapsed_seconds = mean(elapsed_seconds, na.rm = TRUE),
    total_rows = sum(rows, na.rm = TRUE),
    total_split_rows = sum(split_rows_sum, na.rm = TRUE),
    total_train_rows = sum(train_rows, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    total_elapsed_minutes = total_elapsed_seconds / 60,
    total_elapsed_hours = total_elapsed_seconds / 3600,
    mean_elapsed_minutes = mean_elapsed_seconds / 60,
    duration_label = fmt_duration(total_elapsed_seconds)
  ) |>
  sort_stages() |>
  sort_scenarios()

scenario_total_cost <- scenario_stage_cost |>
  group_by(scenario_id, scenario_short, scenario_label) |>
  summarise(
    n_stage_substage_rows = n(),
    stages = paste(sort(unique(as.character(stage_label))), collapse = ", "),
    total_elapsed_seconds = sum(total_elapsed_seconds, na.rm = TRUE),
    training_seconds = sum(total_elapsed_seconds[stage == "training"], na.rm = TRUE),
    inference_seconds = sum(total_elapsed_seconds[stage == "inference"], na.rm = TRUE),
    pairwise_seconds = sum(total_elapsed_seconds[stage == "pairwise_metrics"], na.rm = TRUE),
    total_rows = sum(total_rows, na.rm = TRUE),
    total_split_rows = sum(total_split_rows, na.rm = TRUE),
    total_train_rows = sum(total_train_rows, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    total_elapsed_hours = total_elapsed_seconds / 3600,
    training_hours = training_seconds / 3600,
    inference_hours = inference_seconds / 3600,
    pairwise_hours = pairwise_seconds / 3600,
    share_training = training_seconds / total_elapsed_seconds,
    share_inference = inference_seconds / total_elapsed_seconds,
    share_pairwise = pairwise_seconds / total_elapsed_seconds,
    duration_label = fmt_duration(total_elapsed_seconds)
  ) |>
  sort_scenarios()

write_pipeline_csv(scenario_stage_cost, "09_scenario_stage_cost.csv")
write_pipeline_csv(scenario_total_cost, "10_scenario_total_pipeline_cost.csv")

scenario_total_cost |>
  select(
    scenario_short,
    stages,
    total_elapsed_hours,
    training_hours,
    inference_hours,
    pairwise_hours,
    share_training,
    share_inference,
    share_pairwise,
    total_rows,
    total_split_rows,
    total_train_rows,
    duration_label
  ) |>
  mutate(
    total_elapsed_hours = fmt_number(total_elapsed_hours),
    training_hours = fmt_number(training_hours),
    inference_hours = fmt_number(inference_hours),
    pairwise_hours = fmt_number(pairwise_hours),
    share_training = fmt_pct(share_training),
    share_inference = fmt_pct(share_inference),
    share_pairwise = fmt_pct(share_pairwise),
    total_rows = fmt_number(total_rows, accuracy = 1),
    total_split_rows = fmt_number(total_split_rows, accuracy = 1),
    total_train_rows = fmt_number(total_train_rows, accuracy = 1)
  ) |>
  knitr::kable(
    caption = "Table 9 - Scenario-level pipeline cost from run metadata."
  )
```

# Reliability and error summary

```{r reliability}
error_files <- fs::dir_ls(LOGS_DIR, recurse = TRUE, regexp = "error\\.txt$")

pipeline_errors <- purrr::map_dfr(
  error_files,
  function(path) {
    path_chr <- as.character(path)

    stage <- case_when(
      str_detect(path_chr, "[/\\\\]training[/\\\\]") ~ "training",
      str_detect(path_chr, "[/\\\\]inference[/\\\\]") ~ "inference",
      str_detect(path_chr, "[/\\\\]metrics[/\\\\]") ~ "metrics",
      str_detect(path_chr, "[/\\\\]pairwise_metrics[/\\\\]") ~ "pairwise_metrics",
      str_detect(path_chr, "[/\\\\]statistical_analysis[/\\\\]") ~ "statistical_analysis",
      str_detect(path_chr, "[/\\\\]error_analysis[/\\\\]") ~ "error_analysis",
      TRUE ~ NA_character_
    )

    parts <- stringr::str_split(path_chr, "[/\\\\]", simplify = TRUE)
    scenario_guess <- parts[parts %in% scenario_order][1]
    seed_part <- parts[str_detect(parts, "^seed_")][1]
    seed_guess <- suppressWarnings(as.integer(str_remove(seed_part, "^seed_")))

    error_text <- paste(readLines(path, warn = FALSE, encoding = "UTF-8"), collapse = "\n")

    tibble(
      stage = stage,
      scenario_id = scenario_guess,
      seed = seed_guess,
      error_path = path_chr,
      error_text = error_text,
      error_preview = substr(error_text, 1, 500)
    )
  }
) |>
  add_scenario_info() |>
  sort_stages() |>
  sort_scenarios()

pipeline_error_summary <- pipeline_errors |>
  count(stage, stage_label, scenario_id, scenario_short, seed, name = "n_error_files") |>
  sort_stages() |>
  sort_scenarios()

status_summary <- run_metadata |>
  filter(parse_ok, !is.na(stage)) |>
  count(stage, stage_label, substage, status, name = "n_runs") |>
  sort_stages()

write_pipeline_csv(pipeline_errors, "11_pipeline_error_files.csv")
write_pipeline_csv(pipeline_error_summary, "12_pipeline_error_summary.csv")
write_pipeline_csv(status_summary, "13_run_status_summary.csv")

if (nrow(pipeline_errors) > 0) {
  pipeline_errors |>
    select(stage_label, scenario_short, seed, error_path, error_preview) |>
    knitr::kable(
      caption = "Table 10 - Error files found in pipeline logs."
    )
} else {
  cat("No error.txt files found in the pipeline logs.")
}

status_summary |>
  knitr::kable(
    caption = "Table 11 - Run status summary from metadata files."
  )
```

# Bottleneck analysis

```{r bottlenecks}
bottleneck_table <- unified_stage_cost |>
  arrange(desc(primary_cost_seconds)) |>
  mutate(
    cost_rank = row_number(),
    cumulative_cost_share = cumsum(share_of_primary_cost)
  ) |>
  select(
    cost_rank,
    stage,
    stage_label,
    primary_cost_seconds,
    primary_cost_hours,
    primary_cost_source,
    share_of_primary_cost,
    cumulative_cost_share,
    observed_stage_seconds,
    run_metadata_elapsed_seconds,
    event_elapsed_seconds_sum,
    n_events,
    run_metadata_files
  )

scenario_bottleneck_table <- scenario_total_cost |>
  arrange(desc(total_elapsed_seconds)) |>
  mutate(
    scenario_cost_rank = row_number(),
    scenario_cost_share = total_elapsed_seconds / sum(total_elapsed_seconds, na.rm = TRUE),
    cumulative_scenario_cost_share = cumsum(scenario_cost_share)
  )

write_pipeline_csv(bottleneck_table, "14_stage_bottleneck_table.csv")
write_pipeline_csv(scenario_bottleneck_table, "15_scenario_bottleneck_table.csv")

bottleneck_table |>
  mutate(
    primary_cost_hours = fmt_number(primary_cost_hours),
    share_of_primary_cost = fmt_pct(share_of_primary_cost),
    cumulative_cost_share = fmt_pct(cumulative_cost_share),
    observed_stage_seconds = fmt_number(observed_stage_seconds),
    run_metadata_elapsed_seconds = fmt_number(run_metadata_elapsed_seconds),
    event_elapsed_seconds_sum = fmt_number(event_elapsed_seconds_sum)
  ) |>
  knitr::kable(
    caption = "Table 12 - Stage bottleneck ranking."
  )

scenario_bottleneck_table |>
  select(
    scenario_cost_rank,
    scenario_short,
    total_elapsed_hours,
    training_hours,
    inference_hours,
    pairwise_hours,
    scenario_cost_share,
    cumulative_scenario_cost_share,
    duration_label
  ) |>
  mutate(
    total_elapsed_hours = fmt_number(total_elapsed_hours),
    training_hours = fmt_number(training_hours),
    inference_hours = fmt_number(inference_hours),
    pairwise_hours = fmt_number(pairwise_hours),
    scenario_cost_share = fmt_pct(scenario_cost_share),
    cumulative_scenario_cost_share = fmt_pct(cumulative_scenario_cost_share)
  ) |>
  knitr::kable(
    caption = "Table 13 - Scenario bottleneck ranking."
  )
```

# Figures

## End-to-end stage timeline

```{r fig-stage-timeline, fig.width=11, fig.height=6}
p_stage_timeline <- stage_event_summary |>
  ggplot(aes(y = fct_rev(stage_label), color = stage)) +
  geom_segment(
    aes(x = first_event_utc, xend = last_event_utc, yend = fct_rev(stage_label)),
    linewidth = 5,
    lineend = "round",
    show.legend = FALSE
  ) +
  scale_color_manual(values = stage_palette, drop = FALSE) +
  labs(
    title = "End-to-end pipeline stage timeline",
    subtitle = "Intervals are based on first and last event timestamp for each stage.",
    x = "Time (UTC)",
    y = "Stage"
  ) +
  base_theme()

print(p_stage_timeline)

save_pipeline_figure(
  p_stage_timeline,
  "01_pipeline_stage_timeline.png",
  width = 11,
  height = 6
)
```

## Stage cost share

```{r fig-stage-cost-share, fig.width=10.5, fig.height=6}
p_stage_share <- unified_stage_cost |>
  ggplot(aes(x = fct_reorder(stage_label, primary_cost_seconds), y = primary_cost_hours, fill = stage)) +
  geom_col(width = 0.72, show.legend = FALSE) +
  geom_text(
    aes(label = paste0(round(share_of_primary_cost * 100, 1), "%")),
    hjust = -0.1,
    size = 3.5
  ) +
  coord_flip() +
  scale_fill_manual(values = stage_palette, drop = FALSE) +
  labs(
    title = "Pipeline cost by stage",
    subtitle = "Primary cost uses run metadata when available; otherwise event elapsed/span.",
    x = "Stage",
    y = "Primary cost (hours)"
  ) +
  base_theme()

print(p_stage_share)

save_pipeline_figure(
  p_stage_share,
  "02_pipeline_cost_by_stage.png",
  width = 10.5,
  height = 6
)
```

## Observed stage span vs primary cost

```{r fig-observed-vs-primary, fig.width=10.5, fig.height=6}
p_observed_primary <- unified_stage_cost |>
  select(stage, stage_label, observed_stage_hours, primary_cost_hours) |>
  pivot_longer(
    cols = c(observed_stage_hours, primary_cost_hours),
    names_to = "cost_type",
    values_to = "hours"
  ) |>
  mutate(
    cost_type = recode(
      cost_type,
      observed_stage_hours = "Observed event span",
      primary_cost_hours = "Primary cost"
    )
  ) |>
  ggplot(aes(x = stage_label, y = hours, fill = cost_type)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Observed stage span vs primary cost",
    subtitle = "Observed span is first-to-last event time. Primary cost uses the best available elapsed-time source.",
    x = "Stage",
    y = "Hours",
    fill = "Cost type"
  ) +
  base_theme() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))

print(p_observed_primary)

save_pipeline_figure(
  p_observed_primary,
  "03_observed_span_vs_primary_cost.png",
  width = 10.5,
  height = 6
)
```

## Scenario total cost

```{r fig-scenario-total-cost, fig.width=10.5, fig.height=6}
p_scenario_total <- scenario_total_cost |>
  ggplot(aes(x = scenario_short, y = total_elapsed_hours, fill = scenario_short)) +
  geom_col(width = 0.72, show.legend = FALSE) +
  geom_text(
    aes(label = paste0(round(total_elapsed_hours, 2), " h")),
    vjust = -0.4,
    size = 3.5
  ) +
  scale_fill_manual(values = scenario_palette, drop = FALSE) +
  labs(
    title = "Scenario-level pipeline cost",
    subtitle = "Sum of available run metadata elapsed time by scenario.",
    x = "Scenario",
    y = "Total elapsed time (hours)"
  ) +
  base_theme()

print(p_scenario_total)

save_pipeline_figure(
  p_scenario_total,
  "04_scenario_total_pipeline_cost.png",
  width = 10.5,
  height = 6
)
```

## Scenario cost composition

```{r fig-scenario-composition, fig.width=10.5, fig.height=6}
scenario_composition <- scenario_total_cost |>
  select(scenario_id, scenario_short, training_hours, inference_hours, pairwise_hours) |>
  pivot_longer(
    cols = c(training_hours, inference_hours, pairwise_hours),
    names_to = "cost_component",
    values_to = "hours"
  ) |>
  mutate(
    cost_component = recode(
      cost_component,
      training_hours = "Training",
      inference_hours = "Inference",
      pairwise_hours = "Pairwise metrics"
    )
  )

p_scenario_composition <- scenario_composition |>
  ggplot(aes(x = scenario_short, y = hours, fill = cost_component)) +
  geom_col(width = 0.72) +
  labs(
    title = "Scenario cost composition",
    subtitle = "Training, inference, and pairwise/judge stages from run metadata.",
    x = "Scenario",
    y = "Elapsed time (hours)",
    fill = "Component"
  ) +
  base_theme()

print(p_scenario_composition)

save_pipeline_figure(
  p_scenario_composition,
  "05_scenario_cost_composition.png",
  width = 10.5,
  height = 6
)
```

## Run metadata duration distribution

```{r fig-run-duration-distribution, fig.width=11, fig.height=7}
p_run_distribution <- run_metadata |>
  filter(parse_ok, !is.na(stage), !is.na(elapsed_minutes)) |>
  ggplot(aes(x = substage, y = elapsed_minutes, color = scenario_short)) +
  geom_point(aes(shape = factor(seed)), size = 2.9, position = position_jitter(width = 0.12, height = 0)) +
  facet_wrap(~ stage_label, scales = "free_x") +
  scale_color_manual(values = scenario_palette, drop = FALSE) +
  labs(
    title = "Run metadata duration distribution",
    subtitle = "Each point is one run_metadata.json file.",
    x = "Substage",
    y = "Elapsed time (minutes)",
    color = "Scenario",
    shape = "Seed"
  ) +
  base_theme() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))

print(p_run_distribution)

save_pipeline_figure(
  p_run_distribution,
  "06_run_metadata_duration_distribution.png",
  width = 11,
  height = 7
)
```

## Event count by stage

```{r fig-event-counts, fig.width=10.5, fig.height=6}
p_event_counts <- all_events |>
  count(stage, stage_label, event_type, name = "n_events") |>
  ggplot(aes(x = stage_label, y = n_events, fill = event_type)) +
  geom_col(width = 0.72) +
  labs(
    title = "Event count by stage",
    subtitle = "Stacked by event_type.",
    x = "Stage",
    y = "Number of events",
    fill = "Event type"
  ) +
  base_theme() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))

print(p_event_counts)

save_pipeline_figure(
  p_event_counts,
  "07_event_count_by_stage.png",
  width = 10.5,
  height = 6
)
```

## Rows processed by stage

```{r fig-rows-processed, fig.width=10.5, fig.height=6}
rows_stage_plot <- unified_stage_cost |>
  mutate(
    best_rows = case_when(
      !is.na(run_metadata_split_rows) & run_metadata_split_rows > 0 ~ run_metadata_split_rows,
      !is.na(run_metadata_rows) & run_metadata_rows > 0 ~ run_metadata_rows,
      !is.na(event_rows_sum) & event_rows_sum > 0 ~ event_rows_sum,
      !is.na(num_inference_rows) & num_inference_rows > 0 ~ num_inference_rows,
      !is.na(total_inference_rows) & total_inference_rows > 0 ~ total_inference_rows,
      TRUE ~ NA_real_
    )
  )

p_rows_stage <- rows_stage_plot |>
  filter(!is.na(best_rows)) |>
  ggplot(aes(x = fct_reorder(stage_label, best_rows), y = best_rows, fill = stage)) +
  geom_col(width = 0.72, show.legend = FALSE) +
  coord_flip() +
  scale_fill_manual(values = stage_palette, drop = FALSE) +
  scale_y_continuous(labels = scales::comma_format()) +
  labs(
    title = "Rows processed or reported by stage",
    subtitle = "Uses best available row count from metadata, events, or summaries.",
    x = "Stage",
    y = "Rows"
  ) +
  base_theme()

print(p_rows_stage)

save_pipeline_figure(
  p_rows_stage,
  "08_rows_processed_by_stage.png",
  width = 10.5,
  height = 6
)
```

## Error files by stage

```{r fig-errors, fig.width=9.5, fig.height=5.5}
if (nrow(pipeline_error_summary) > 0) {
  p_errors <- pipeline_error_summary |>
    group_by(stage, stage_label) |>
    summarise(n_error_files = sum(n_error_files), .groups = "drop") |>
    ggplot(aes(x = stage_label, y = n_error_files, fill = stage)) +
    geom_col(width = 0.72, show.legend = FALSE) +
    scale_fill_manual(values = stage_palette, drop = FALSE) +
    labs(
      title = "Error files by stage",
      subtitle = "Counts error.txt files discovered under logs.",
      x = "Stage",
      y = "Error files"
    ) +
    base_theme() +
    theme(axis.text.x = element_text(angle = 25, hjust = 1))

  print(p_errors)

  save_pipeline_figure(
    p_errors,
    "09_error_files_by_stage.png",
    width = 9.5,
    height = 5.5
  )
}
```

# Paper-ready table

```{r paper-ready}
paper_pipeline_table <- unified_stage_cost |>
  select(
    stage_label,
    notebook,
    n_events,
    run_metadata_files,
    observed_stage_seconds,
    primary_cost_seconds,
    primary_cost_source,
    share_of_primary_cost,
    event_rows_sum,
    run_metadata_rows,
    run_metadata_split_rows,
    run_metadata_train_rows
  ) |>
  mutate(
    observed_stage_hours = fmt_number(observed_stage_seconds / 3600),
    primary_cost_hours = fmt_number(primary_cost_seconds / 3600),
    share_of_primary_cost = fmt_pct(share_of_primary_cost),
    event_rows_sum = fmt_number(event_rows_sum, accuracy = 1),
    run_metadata_rows = fmt_number(run_metadata_rows, accuracy = 1),
    run_metadata_split_rows = fmt_number(run_metadata_split_rows, accuracy = 1),
    run_metadata_train_rows = fmt_number(run_metadata_train_rows, accuracy = 1)
  ) |>
  select(
    stage_label,
    notebook,
    n_events,
    run_metadata_files,
    observed_stage_hours,
    primary_cost_hours,
    primary_cost_source,
    share_of_primary_cost,
    event_rows_sum,
    run_metadata_rows,
    run_metadata_split_rows,
    run_metadata_train_rows
  )

write_pipeline_csv(paper_pipeline_table, "16_paper_ready_pipeline_runtime_table.csv")

paper_pipeline_table |>
  knitr::kable(
    caption = "Table 14 - Paper-ready pipeline runtime table."
  )
```

# Interpretation notes

```{r methodological-notes, results='asis'}
cat("
## How to interpret this report

This report uses two notions of time.

Observed stage span is the time between the first and last event timestamp in each stage log. It is useful for timeline analysis.

Primary cost is the best available elapsed-time estimate for each stage. When run_metadata.json files exist, the primary cost is the sum of run metadata elapsed_seconds. Otherwise, the report falls back to elapsed_seconds found in event rows, and then to the event timestamp span.

Because some stages include multiple scenario/seed runs, primary cost can differ from wall-clock duration. It estimates accumulated compute or task time, while observed stage span estimates calendar time.

Training and inference are model-execution stages. Metrics, statistical analysis, and error analysis are mostly post-processing stages. Pairwise metrics can include both additional inference-style injected-only runs and judge-based win-rate runs.

C0 and C1 have no adapter-training cost, but they can have inference, injected-only, and pairwise metric costs.
")
```

# Generated files

Tables are saved in:

```text
tables/pipeline_runtime_analysis/
```

Figures are saved in:

```text
figures/pipeline_runtime_analysis/
```

Main files:

```text
01_stage_paths.csv
02_all_pipeline_events.csv
03_stage_event_timeline_summary.csv
04_stage_summary_json_records.csv
05_all_run_metadata.csv
06_stage_cost_from_run_metadata.csv
07_unified_stage_cost_summary.csv
08_pipeline_total_summary.csv
09_scenario_stage_cost.csv
10_scenario_total_pipeline_cost.csv
11_pipeline_error_files.csv
12_pipeline_error_summary.csv
13_run_status_summary.csv
14_stage_bottleneck_table.csv
15_scenario_bottleneck_table.csv
16_paper_ready_pipeline_runtime_table.csv

01_pipeline_stage_timeline.png
02_pipeline_cost_by_stage.png
03_observed_span_vs_primary_cost.png
04_scenario_total_pipeline_cost.png
05_scenario_cost_composition.png
06_run_metadata_duration_distribution.png
07_event_count_by_stage.png
08_rows_processed_by_stage.png
09_error_files_by_stage.png
```
